### 2021-01-06 summary.Rmd
typo

---
title: "Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Clustering and Parameter Estimation"
author: John Koo, Minh Tang, Michael Trosset
output:
  pdf_document:
    citation_package: natbib
    number_sections: yes
# output: html_document
fontsize: 12pt
# geometry: "left=1cm,right=1cm,top=1cm,bottom=1.5cm"
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square}
- \usepackage{verbatim}
bibliography: misc.bib
abstract: |
  In this paper, we connect two random graph models, the Popularity 
  Adjusted Block Model (PABM) and the Generalized Random Dot Product Graph 
  (GRDPG) and use properties established in this connection to aid in 
  community detection and parameter estimation. In particular, we note that the
  PABM can be represented as latent positions such that points within the same 
  community lie on a subspace, and the subspaces that represent each 
  community are orthogonal to one another. Using this property as well as the 
  asymptotic properties of Adjacency Spectral Embedding (ASE) of the GRDPG, we 
  are able to establish theoretical asymptotic results of our community 
  detection and parameter estimation methods for the PABM. 
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      # eval = FALSE,
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 3, 
                      fig.width = 5, 
                      fig.dpi = 300)

options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
```

```{r}
import::from(magrittr, `%>%`)
library(plot.matrix)
library(ggplot2)
library(mclust)

theme_set(theme_bw())
```

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{indep}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}

# Introduction

Statistical analysis on graphs or networks often involves the partitioning of 
a graph into disconnected subgraphs or clusters. This is often motivated by the 
belief that there exist underlying and unobserved communities to which each 
vertex of the graph belongs, and edges between pairs of vertices are determined 
by drawing from a probability distribution based on the community relationships 
between each pair. The goal of this analysis then is population community 
detection, or the recovery of the true underlying community labels for each 
vertex, up to permutation (with some additional parameter estimation being of 
possible interest), assuming some random graph model. One such model is 
the Stochastic Block Model (SBM), first proposed by 
\citet{doi:10.1080/0022250X.1971.9989788}, which assumes that the edge 
probability from one vertex to another follows a Beronulli distribution with 
fixed probabilities for each pair of community labels. Other random 
graph models have been proposed and studied, such as the Degree-Corrected Block 
Model (DCBM), introduced by \citet{Karrer_2011} and the Mixed Membership 
Stochastic Block Model (MMSBM), introduced by  \citet*{NIPS2008_3578}. The 
Popularity Adjusted Block Model (PABM) was then introduced by 
\citet*{307cbeb9b1be48299388437423d94bf1} as a generalization of the SBM to 
address the heterogeneity of edge probabilities within and between communities 
while still maintaining distinct community structure.

The underlying similarity among the SBM, PABM, and other such models is that 
they involve a symmetric edge probability matrix $P \in [0, 1]^{n \times n}$ 
where $n$ is the number of vertices in the graph. An undirected and unweighted 
graph is then drawn from this edge probability matrix such that the existence 
of an edge between each pair of vertices $i$ and $j$ is given by 
$\Bernoulli(P_{ij})$. For example, for the SBM with two communities for which 
the within-community edge probability is $\xi$ and the between-community edge 
probability is $\eta$, the entries of $P$ consist of $\xi$ and $\eta$.

The Random Dot Product Graph (RDPG) model proposed by 
\citet*{10.1007/978-3-540-77004-6_11} is another graph model with Bernoulli edge 
probabilities. Under this model, each vertex of the graph can be represented by 
a point in a latent Euclidean space such that the edge probability between any 
pair of vertices is given by their corresponding dot product in the latent 
space, i.e., given a latent positions $x_1, ..., x_n \in \mathbb{R}^d$, the 
edge probability matrix is $P = X X^\top$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$. The SBM is 
equivalent to a special case of the RDPG model in which all vertices of a 
given community share the same position in the latent space 
\cite{lyzinski2014}. It has also been shown that similar 
random graph models, including both the MMSBM and DCBM, can be 
represented in this way \cite{rubindelanchy2017consistency} 
\cite{lyzinski2014}. An analogous 
property exists for the PABM but not for the RDPG model but under the 
*Generalized* Random Dot Product Graph (GRDPG) model. This relationship will be 
explored in this paper and exploited to construct algorithms for community 
detection and parameter estimation for the PABM.

In this paper, we will only consider undirected graphs, that is the edge weight 
from vertex $i$ to vertex $j$ is equal to the edge weight in the opposite 
direction, from vertex $j$ to vertex $i$. Furthermore, we will only consider 
unweighted graphs with binary ($0, 1$) edge weights We will also assume that 
graphs are hollow, i.e., there are no edges from a vertex to itself. All such 
graphs can be represented by a symmetric adjacency matrix 
$A \in \{0, 1\}^{n \times n}$ for which $A_{ij} = 1$ if there exists an edge 
between vertices $i$ and $j$ and $0$ otherwise, and $A$ is an element-wise 
independent Bernoulli draw from a symmetric edge probability matrix 
$P \in [0, 1]^{n \times n}$.

# Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph

## The popularity adjusted block model (PABM) and the generalized random dot product graph

**Definition 1** (Popularity Adjusted Block Model) 
\cite{307cbeb9b1be48299388437423d94bf1}. 
Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
set of $n$ 
vertices, $V$. Each vertex has a community label $1, ..., K$, and the rows and 
columns of $P$ are arranged by community label such that $n_k \times n_l$ block 
$P^{(kl)}$ describes the edge probabilities between vertices in communities 
$k$ and $l$ ($P^{(lk)} = (P^{(kl)})^\top$). 
Let graph $G = (V, E)$ be an undirected, unweighted graph such 
that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
realization of $\Bernoulli(P)$, i.e., 
$A_{ij} \indep \Bernoulli(P_{ij})$ for $i > j$ 
($A_{ij} = A_{ji}$ and $A_{ii} = 0$).

If each block $P^{(kl)}$ can be written as the outer product of two vectors:

\begin{equation}
  P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top
\end{equation}

for a set of $K^2$ fixed vectors $\{\lambda^{(st)}\}_{s, t = 1}^K$ where each 
$\lambda^{(st)}$ is a column vector 
of dimension $n_s$, then graph $G$ and its corresponding adjacency matrix $A$ 
is a realization of a popularity adjusted block model with parameters 
$\{\lambda^{(st)}\}_{s, t = 1}^K$.

We will use the notation $A \sim PABM(\{\lambda^{(kl)}\}_K)$ to denote a random 
adjacency matrix $A$ drawn from a PABM with parameters $\lambda^{(kl)}$ 
consisting of $K$ underlying communities.

**Definition 2** (Generalized Random Dot Product Graph) 
\cite{rubindelanchy2017statistical}.
Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
set of $n$ vertices, $V$. If $\exists X \in \mathbb{R}^{n \times d}$ such that

\begin{equation}
  P = X I_{pq} X^\top
\end{equation}

for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then 
graph $G = (V, E)$ with adjacency matrix $A$ such that 
$A_{ij} \indep \Bernoulli(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and 
$A_{ii} = 0$) is a draw from the generalized random dot product graph model 
with latent positions $X$ and signature $(p, q)$. More precisely, if vertices 
$i$ and $j$ have latent positions $x_i$ and $x_j$ respectively, then the edge 
probability between the two is $P_{ij} = x_i^\top I_{pq} x_j$, and $X$ contains 
the latent positions as rows $x_i^\top$.

We will use the notation $A \sim GRDPG_{p,q}(X)$ to denote a random adjacency 
matrix $A$ drawn from latent positions $X$ and signature $(p, q)$.

**Definition 3**. The indefinite orthogonal group with signature $(p, q)$ is 
the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{pq} Q^{\top} = I_{pq}\}$, 
denoted as $\mathbb{O}(p, q)$ \cite{rubindelanchy2017statistical}.

**Remark**. 
Like the RDPG, the latent positions of a GRDPG are not unique 
\cite{rubindelanchy2017statistical}. 
More specifically, if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
$Q \in \mathbb{O}(p, q)$, 
$(Q x_i)^\top I_{pq} (Q x_j) = x_i^\top (Q^\top I_{pq} Q) x_j = 
x_i^\top I_{pq} x_j = P_{ij}$. 
Unlike in the RDPG case, transforming the latent positions via multiplication 
by $Q \in \mathbb{O}(p, q)$ does not necessarily maintain interpoint angles or 
distances.

## Connecting the PABM to the GRDPG

**Theorem 1** (Connecting the PABM to the GRDPG for $K = 2$). 
Let

$$X = \begin{bmatrix} 
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\ 
0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
\end{bmatrix}$$

$$U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
0 & 1 & 0 & 0 \end{bmatrix}$$

as in Definition 1. 
Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_2)$ are 
equivalent.

*Proof*. 
This is given by straightforward matrix multiplication. It suffices to show 
that

$$X U I_{3, 1} U^\top X^\top = 
\begin{bmatrix} 
  \lambda^{(11)} (\lambda^{(11)})^\top & \lambda^{(12)} (\lambda^{(21)})^\top \\
  \lambda^{(21)} (\lambda^{(12)})^\top & \lambda^{(22)} (\lambda^{(22)})^\top
\end{bmatrix}$$

**Remark**. While we can just perform the matrix multiplication to show the 
equivalence, it is more illustrative to look at a few intermediate steps. Note 
that the product of the three inner matrices results in a permutation matrix 
with fixed points at positions $1$ and $4$ and a cycle of order 2 swappoing 
positions $2$ and $3$:

$$U I_{3, 1} U^\top = \begin{bmatrix} 
  1 & 0 & 0 & 0 \\ 
  0 & 0 & 1 & 0 \\ 
  0 & 1 & 0 & 0 \\ 
  0 & 0 & 0 & 1 
\end{bmatrix} = \Pi$$

Since $U$ is orthonormal and $I_{3, 1}$ is diagonal, $\Pi = U I_{3, 1} U^\top$ 
is a spectral decomposition of this permutation matrix. Note that the two fixed 
points result in eigenvalues of $+1$ with corresponding eigenvectors $e_i$ 
where $i = 1, 4$ corresponding to the locations of the fixed points, and the 
cycle of order two results in two eigenvalues $\pm 1$ with corresponding 
eigenvectors $(e_i \pm e_j) / \sqrt{2}$ where $i = 2, j = 3$, pair that is 
swapped.

**Theorem 2** (Generalization to $K > 2$). There exists a block diagonal matrix 
$X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters 
$\{\lambda^{(kl)}\}_K$ and $U \in \mathbb{R}^{K^2 \times K^2}$ that is fixed 
for each $K$ such that $A \sim GRDPG_{K (K+1) / 2, K (K-1) / 2}(XU)$ and 
$A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.

*Proof*. 
Let 
$$\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}$$

\begin{equation}
X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
\end{equation}

$$L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
\mathbb{R}^{n \times K}$$

$$Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
\mathbb{R}^{n \times K^2}$$

Then $P = X Y^\top$.

Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
$\Pi$, resulting in $P = X \Pi X^\top$.  
The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.

Then $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
eigenvalues equal to $-1$. $\Pi$ has the decomposed form

\begin{equation}
\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
\end{equation}

The edge probability matrix then can be written as:

\begin{equation}
P = X U I_{p, q} (X U)^\top
\end{equation}

\begin{equation}
p = K (K + 1) / 2
\end{equation}

\begin{equation}
q = K (K - 1) / 2
\end{equation}

and we can describe the PABM with $K$ communities as a GRDPG with latent 
positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.

**Example** ($K = 3$). Using the same notation as before:

$$X = \begin{bmatrix} 
\lambda^{(11)} & \lambda^{(12)} & \lambda^{(13)} & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & \lambda^{(21)} & \lambda^{(22)} & \lambda^{(23)} & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & \lambda^{(31)} & \lambda^{(32)} & \lambda^{(33)}
\end{bmatrix}$$

$$Y = \begin{bmatrix} 
\lambda^{(11)} & 0 & 0 & \lambda^{(12)} & 0 & 0 & \lambda^{(13)} & 0 & 0 \\
0 & \lambda^{(21)} & 0 & 0 & \lambda^{(22)} & 0 & 0 & \lambda^{(23)} & 0 \\
0 & 0 & \lambda^{(31)} & 0 & 0 & \lambda^{(32)} & 0 & 0 & \lambda^{(33)}
\end{bmatrix}$$

Then $P = X Y^\top$ and $Y = X \Pi$ where $\Pi$ is a permutation matrix 
consisting of $3$ fixed points and $3$ cycles of order 2:

$$\Pi = \begin{bmatrix} 
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
\end{bmatrix}$$

* Positions 1, 5, 9 are fixed.
* The cycles of order 2 are $(2, 4)$, $(3, 7)$, and $(6, 8)$.

Therefore, we can decompose $\Pi = U I_{6, 3} U^\top$ where the first three 
columns of $U$ consist of $e_1$, $e_5$, and $e_9$ corresponding to the fixed 
positions $1$, $5$, and $9$, the next three columns consist of eigenvectors 
$(e_k + e_l) / \sqrt{2}$, and the last three columns consist of eigenvectors 
$(e_k - e_l) / \sqrt{2}$, where pairs $(k, l)$ correspond to the cycles of 
order 2 described above.

The latent positions are the rows of  
$$XU = \begin{bmatrix}
  \lambda^{(11)} & 0 & 0 & 
  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 & 
  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 \\
  0 & \lambda^{(22)} & 0 & 
  \lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} & 
  -\lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} \\
  0 & 0 & \lambda^{(33)} & 
  0 & \lambda^{(31)} / \sqrt{2} & \lambda^{(32)} / \sqrt{2} & 
  0 & -\lambda^{(31)} / \sqrt{2} & -\lambda^{(32)} / \sqrt{2}
\end{bmatrix}$$

# Methods

Two inference objectives arise from the PABM:

1. Community membership identification (up to permutation).
2. Parameter estimation (estimating $\lambda^{(kl)}$'s).

In our methods, we assume that $K$, the number of communities, is known 
beforehand and does not require estimation.

## Related work

\citet*{307cbeb9b1be48299388437423d94bf1}, who first proposed the PABM, used 
modularity maximization for clustering and parameter estimation. They were able 
to show that as the sample size increases, the proportion of misclassified 
community labels (up to permutation) goes to 0. However, because modularity 
maximization is NP-hard, the authors of this paper used an approximation 
based on the Extreme Points algorithm \cite{le2016}.

\citet*{noroozi2019estimation} proposed using sparse subspace clustering 
to identify the community memberships given either an edge probability matrix 
$P$ or an adjacency matrix $A$. In the case that $P$ is known, the community 
memberships can be identified exactly (up to permutation). This is supported by 
our Theorem 2, which shows that the ASE of edge probability matrix $P$ consists 
of $K$ $K$-dimensional subspaces lying in $\mathbb{R}^{K^2}$. A similar 
procedure can be applied if $P$ is unknown and we have an observation $A$, but 
the theoretical guarantees of this method applied to the PABM are unknown.

## Community detection

### Using edge probability matrix $P$

We previously stated one possible set of latent positions that result in the 
edge probability matrix of a PABM, $P = (XU) I_{pq} (XU)^\top$. If we have 
(or can estimate) $XU$ directly, then both the community detection and parameter 
identification problem are trivial since $U$ is orthonormal and fixed for each 
value of $K$. However, direct identification or estimation of $XU$ is not 
possible \cite{rubindelanchy2017statistical}.

If we decompose $P = Z I_{pq} Z^\top$, then
$\exists Q \in \mathbb{O}(p, q)$ such that $XU = Z Q$. Even if we start with 
the exact edge probability matrix, we cannot recover the "original" latent 
positions $XU$. Note that unlike in the case of the RDPG, 
$Q$ is not an orthogonal matrix. If $z_i$'s are the rows of $XU$, then 
$||z_i - z_j||^2 \neq ||Q z_i - Q z_j||^2$, and 
$\langle z_i, z_j \rangle \neq \langle Q z_i, Q z_j \rangle$. This prevents 
us from using the properties of $XU$ directly. In particular, if 
$Q \in \mathbb{O}(n)$, then we could use the fact that 
$\langle z_i, z_j \rangle = \langle Q z_i, Q z_j \rangle = 0$ if vertices $i$ 
and $j$ are in different communities.

We can note from the explicit form of $XU$ that it represents points in 
$\mathbb{R}^{K^2}$ such that points within each community lie on $K$-dimensional 
subspaces. Furthermore, the subspaces are orthogonal to each other. 
Multiplication by $Q \in \mathbb{O}(p, q)$ removes the orthogonality property 
but retains the property that each community is represented by a $K$-dimensional 
subspace. Using this property, previous work proposes the use of subspace 
clustering \cite{noroozi2019estimation} \cite{soltanolkotabi2014}.

**Lemma 1.** 
Let $P = V D V^\top$ be the spectral decomposition of the edge probability 
matrix for a PABM. Then $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is 
defined as in equation (3).

*Proof*. 
By Theorem 2, $P = X U I_{p, q} U^\top X^\top$, where $X$ is defined as in equation (3) and $p$ and $q$ are defined as in equations (6) and (7). Alternatively, the spectral decomposition can be written as $P = V D V^\top = V |D|^{1/2} I_{p, q} |D|^{1/2} V^\top$ for the same $(p, q)$ and $|\cdot|^{1/2}$ is applied entry-wise. Thus for some $Q \in \mathbb{O}(p, q)$,

$$X U Q = V |D|^{1/2}$$

Therefore, using the fact that $U U^\top = I$ and $V^\top V = I$,

$$(V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top = 
(X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top$$

The righthand side becomes

$$\begin{split}
  (X U Q) ((X U Q)^\top (X U Q))^{-1} (X U Q)^\top & 
    = X U Q Q^{-1} U^\top (X^\top X)^{-1} U (Q^\top)^{-1} Q^\top U^\top X^\top \\
  & = X U U^\top (X^\top X)^{-1} U U^\top X^\top \\
  & = X (X^\top X)^{-1} X^\top
\end{split}$$

The lefthand side becomes:

$$\begin{split}
  (V |D|^{1/2}) ((V |D|^{1/2})^\top (V |D|^{1/2}))^{-1} (V |D|^{1/2})^\top & =
    V |D|^{1/2} |D|^{-1/2} (V^\top V)^{-1} |D|^{-1/2} |D|^{1/2} V^\top \\
  & = V V^\top
\end{split}$$

**Theorem 3.** 
Let $P = V D V^\top$ be the spectral decomposition of the edge probability
matrix. Let $B = V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
of different communities.

*Proof*.
By Lemma 1, $V V^\top = X (X^\top X)^{-1} X^\top$ where $X$ is defined as 
in Theorem 2. Since $X$ is block diagonal with each block corresponding 
to one community, $X (X^\top X)^{-1} X^\top$ is also a block diagonal matrix 
with each block corresponding to a community and zeros elsewhere. Therefore, 
if vertices $i$ and $j$ belong to different communities, then the $ij$^th^ 
element of $X (X^\top X)^{-1} X^\top = V V^\top = B$ is 0.

\begin{algorithm}[H]
  \DontPrintSemicolon
  \SetAlgoLined
  \KwData{Edge probability matrix $P$, number of communities $K$}
  \KwResult{Community labels $1, ..., K$}
    Compute the spectral decomposition $P = V D V^\top$.\;
    Compute the inner product matrix $B = |V V^\top|$, applying $|\cdot|$ 
    entry-wise.\;
    Construct graph $G$ using $B$ as edge similarities.\;
    Identify the connected components of $G$ and map each to community labels 
    $1, ..., K.$\;
  \caption{PABM clustering on the edge probability matrix.}
\end{algorithm}

### Using adjacency matrix $A$

The adjacency embedding of $A$ approaches latent positions that form $P$ as 
the number of vertices $n$ increases. More 
precisely, let $\{\lambda^{(kl)}\}_K \sim \mathcal{F}_K$ for some joint 
distribution consisting of $K$ underlying communities $\mathcal{F}_K$. Then the 
latent positions $XU \sim \mathcal{G}_K$ for some related joint distribution 
with $K$ underlying communities $\mathcal{G}_K$. Denote $Z_n$ as a sample of size 
$n$ from $\mathcal{G}_K$ and adjacency matrix $A_n$ as one draw from edge 
probability matrix $P_n = Z_n I_{pq} Z_n^\top$. Let $\hat{Z}_n$ be the 
adjacency embedding of $A_n$ with rows $(\hat{z}_i^{(n)})^\top$. Then by
\citet{rubindelanchy2017statistical},

\begin{equation}
  \max\limits_{i \in \{1, ..., n\}} ||Q_n \hat{z}_i^{(n)} - z_i^{(n)}|| = 
  O_P\bigg(\frac{(\log n)^c}{n^{1/2}} \bigg)
\end{equation}

for some $c>0$ and sequence of $Q_n \in \mathbb{O}(p, q)$. In addition, 
\citeauthor{rubindelanchy2017statistical} produce a central limit theorem 
result.

**Theorem 4**. Let $\hat{V}^{(n)} \in \mathbb{R}^{n \times K^2}$ be the matrix 
of $K^2$ eigenvectors of $A_n$ corresponding to the $K (K+1) / 2$ most positive 
eigenvalues and $K (K-1) / 2$ most negative eigenvalues with rows
$(\hat{v}_i^{(n)})^\top$. Let $(i, j)$ correspond to pairs belonging to 
different communities. Then for some $c > 0$,

\begin{equation}
  \max\limits_{i, j} ||(\hat{v}^{(n)}_i)^\top \hat{v}_j^{(n)}|| = 
    O_P\bigg( \frac{(\log n)^c}{n \sqrt{\rho_n}}\bigg)
\end{equation}

where $\lim\limits_{n \to 0} \rho_n = 0$ is a sparsity parameter introduced 
by \citet{307cbeb9b1be48299388437423d94bf1}. More precisely, 
$\rho_n = o(\log n / \sqrt{n})$ \cite{rubindelanchy2017statistical}.

This leads to the following algorithm:

\begin{algorithm}[H]
  \DontPrintSemicolon
  \SetAlgoLined
  \KwData{Adjacency matrix $A$, number of communities $K$}
  \KwResult{Community assignments $1, ..., K$}
    Compute the eigenvectors of $A$ that correspond to the $K (K+1) / 2$ most 
    positive eigenvalues and $K (K-1) / 2$ most negative eigenvalues. Construct 
    $V$ using these eigenvectors as its columns.\;
    Compute $B = |V V^\top|$, applying $|\cdot|$ entry-wise.\;
    Construct graph $G$ using $B$ as its similarity matrix.\;
    Partition $G$ into $K$ disconnected components 
    (e.g., using edge thresholding or spectral clustering).\;
    Map each partition to the community labels $1, ..., K$.\;
  \caption{PABM clustering on the adjacency matrix.}
\end{algorithm}

Theorem 4 implies that as $n \to \infty$, the number (not just proportion) of 
misclassified vertices, up to permutation, outputted by algorithm 2 goes to 0.

## Parameter estimation

For any $P$ edge probability matrix for the PABM such that the rows and 
columns are organized by community, the $kl$^th^ block is an outer product 
of two vectors, i.e., $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. 
Therefore, given $P^{(kl)}$, $\lambda%{(kl)}$ and $\lambda^{(lk)}$ are 
solvable exactly (up to multiplication by $-1$) using singular value 
decomposition. More specifically, let $P = \sigma^2 u v^\top$ be the 
singular value decomposition of $P$. $u \in \mathbb{R}^{n_k}$ and 
$v \in \mathbb{R}^{n_l}$ are vectors and 
$\sigma^2 > 0$ is a scalar. Then $\lambda^{(kl)} = \pm \sigma u$ and 
$\lambda^{(lk)} = \pm \sigma v$.

\begin{algorithm}[H]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{PABM parameter estimation using the edge probability matrix.}
  \KwData{Edge probability matrix $P$, community assignments $1, ..., K$.}
  \KwResult{PABM parameters $\{\lambda^{(kl)}\}_K$}
  Arrange the rows and columns of $P$ by community such that each $P^{(kl)}$ 
  block consists of edge probabilities between communities $k$ and $l$.\; 
  \For{$k, l = 1, ..., K$, $k \geq l$} {
    Compute $P^{(kl)} = (\sigma^{(kl)})^2 u^{(kl)} (v^{(kl)})^\top$, the SVD of 
    the $kl^{th}$ block.\;
    Assign $\lambda^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
    $\lambda^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
  }
\end{algorithm}

Given the adjacency matrix $A$ instead of edge probability matrix $P$, we can 
simply use plug-in estimators:

\begin{algorithm}[H]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{PABM parameter estimation using the adjacency matrix.}
  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(kl)}\}_K$.}
  Arrange the rows and columns of $A$ by community such that each 
  $A^{(kl)}$ block consists of estimated edge probabilities between 
  communities $k$ and $l$.\;
  \For{$k, l = 1, ..., K$, $k \leq l$} {
    Compute $A^{(kl)} = U \Sigma V^\top$, the SVD of the $kl^{th}$ 
    block.\;
    Assign $u^{(kl)}$ and $v^{(kl)}$ as the first columns of $U$ and $V$. 
    Assign $(\sigma^{(kl)})^2 \leftarrow \Sigma_{11}$.\;
    Assign $\hat{\lambda}^{(kl)} \leftarrow \pm \sigma^{(kl)} u^{(kl)}$ and 
    $\hat{\lambda}^{(lk)} \leftarrow \pm \sigma^{(kl)} v^{(kl)}$.
  }
\end{algorithm}

**Theorem 5**. Under regularity and sparsity assumptions, and under the further 
assumption that $K$ is fixed and community labels are known,

\begin{equation}
\max_{k, l \in \{1, ..., K\}} 
||\hat{\lambda}^{(kl)} - \lambda^{(kl)}|| = 
O_P \bigg(\frac{(\log n_k)^c}{\sqrt{n_k}} \bigg)
\end{equation}

*Proof*. 
Let $P$ and $A$ be organized by community such that the elements of blocks 
$P^{(kl)}$ and $A^{(kl)}$ correspond to the edges between communities $k$ and 
$l$.

*Case $k = l$*. $P^{(kk)}$ and $A^{(kk)}$ represent within-community edge 
probabilities and edges for community $k$.  
By definition, 
$P^{(kk)} = \lambda^{(kk)} (\lambda^{(kk)})^\top$. This implies that the 
signular value decomposition 
$P^{(kk)} = \sigma_{kk}^2 u^{(kk)} (u^{(kk)})^\top$ has one singular value and 
one pair of singular vectors ($P^{(kk)}$ is symmetric, so the left and right 
singular vectors are identical). Then $\lambda^{(kk)} = \sigma_{kk} u^{(kk)}$.  
Let $\hat{U}^{(kk)} \hat{\Sigma}^{(kk)} (\hat{U}^{(kk)})^\top$ be the singular 
value decomposition of $A^{(kk)}$, and let 
$\hat{\sigma}_{kk}^2 \hat{u}^{(kk)} (\hat{u}^{(kk)})^\top$ be its 
one-dimensional approximation. Define 
$\hat{\lambda}^{(kk)} = \hat{\sigma}_{kk} \hat{u}^{(kk)}$. Then 
$\hat{\lambda}^{(kk)}$ is the adjacency spectral embedding approximation of 
$\lambda^{(kk)}$.  
Then by Theorem 5 from \citet{rubindelanchy2017statistical}, the adjacency 
spectral embedding $\hat{\lambda}^{(kk)}$ approximates $\lambda^{(kk)}$ at rate 
$\frac{(\log n_k)^c}{\sqrt{n_k}}$.

*Case $k \neq l$*. $P^{(kl)}$ and $A^{(kl)}$ represent edge probabilities and 
edges between communities $k$ and $l$. Note that $P^{(kl)} = (P^{(lk)})^\top$.  
By definition, $P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$. As in the 
$k = l$ case, we note that the singular value decomposition 
$P^{(kl)} = \sigma_{kl}^2 u^{(kl)} (v^{(kl)})^\top$ is one-dimensional and 
$\lambda^{(kl)} = \sigma_{kl} u^{(kl)}$. (We can also note that the SVD of 
$P^{(lk)} = \sigma_{kl}^2 v^{(kl)} (u^{(kl)})^\top$, i.e., 
$\sigma_{kl} = \sigma_{lk}$, $u^{(kl)} = v^{(lk)}$, and 
$v^{(kl)} = u^{(lk)}$.)  
Now consider the Hermitian dilation

$$M^{(kl)} = 2 \begin{bmatrix} 0 & P^{(kl)} \\ P^{(lk)} & 0 \end{bmatrix}$$

which is a symmetric $(n_k + n_l) \times (n_k + n_l)$ matrix. It can be shown 
that the eigendecomposition of $M^{(kl)}$ is

$$M^{(kl)} = 
\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix} \times 
\begin{bmatrix} \sigma^2_{kl} & 0 \\ 0 & -\sigma^2_{kl} \end{bmatrix} \times
\begin{bmatrix} u^{(kl)} & -u^{(kl)} \\ v^{(kl)} & v^{(kl)} \end{bmatrix}^\top$$

Thus treating $M^{(kl)}$ as the edge probability matrix of a GRDPG, we have 
latent positions in $\mathbb{R}^2$ given by

$$\begin{bmatrix} 
  \sigma_{kl} u^{(kl)} & \sigma_{kl} u^{(kl)} \\ 
  \sigma_{kl} v^{(kl)} & -\sigma_{kl} v^{(kl)} 
\end{bmatrix} = 
\begin{bmatrix} 
  \lambda^{(kl)} & \lambda^{(kl)} \\ 
  \lambda^{(lk)} & -\lambda^{(lk)} 
\end{bmatrix}$$

Now consider

$$\hat{M}^{(kl)} = \begin{bmatrix} 0 & A^{(kl)} \\ A^{(lk)} & 0 \end{bmatrix}$$ 
Then $\hat{M}^{(kl)} = M^{(kl)} + E'$ where

$$E' = \begin{bmatrix} 0 & E \\ E^\top & 0 \end{bmatrix}$$

and $E$ is the 
$n_k \times n_l$ matrix of independent noise (to generate the Bernoulli entries 
in $A^{(kl)}$.  
Then $\hat{M}^{(kl)}$ is an adjacency matrix drawn from $M^{(kl)}$, so its adjacency 
spectral embedding, given by

$$\begin{bmatrix} 
  \hat{\lambda}^{(kl)} & \hat{\lambda}^{(kl)} \\ 
  \hat{\lambda}^{(lk)} & -\hat{\lambda}^{(lk)} 
\end{bmatrix}$$

where each $\hat{\lambda}^{(kl)}$ is defined as in Algorithm 4, approximates 
the latent positions of $M^{(kl)}$ up to indefinite orthogonal transformation 
by the rate given in Theorem 5 of \citet{rubindelanchy2017statistical}.  
In this case, the indefinite orthogonal transformation $W_*$ in the GRDPG 
result \cite{rubindelanchy2017statistical} is of the form 
$U^\top \hat{U}$. The eigenvalues of $M$ are distinct since the signature 
for this GRDPG is $(1, 1)$, and $U^\top \hat{U}$ is block diagonal, 
resulting in $W_* \stackrel{P}{\to} I$. Therefore, the adjacency spectral embedding of $\hat{M}^{(kl)}$ is 
a direct estimation of the specific latent positions outlined for $M^{(kl)}$, 
up to sign flip.

# Simulated Examples

For each simulation, the popularity vectors $\{\lambda^{(kl)}\}_K$ are drawn 
from a joint distribution, the edge probability matrix $P$ is constructed using 
the popularity vectors, and finally an unweighted and undirected adjacency 
matrix $A$ is drawn from $P$. Algorithm 2 is then used for community detection, 
and this method is compared against sparse subspace clustering 
\cite{soltanolkotabi2014} \cite{noroozi2019estimation} and modularity 
maximization \cite{307cbeb9b1be48299388437423d94bf1} \cite{igraph}. True 
community labels are used with Algorithm 4 to estimate the popularity 
vectors, and this method is then compared against an MLE-based estimator 
described in \citet{noroozi2019estimation} and
\citet{307cbeb9b1be48299388437423d94bf1}.

Sparse subspace clustering is described as follows. First, an embedding of the 
affinity matrix $A$ is required, and for this, we use spectral decomposition 
using the first $K (K + 1) / 2$ and last $K (K - 1) / 2$ eigenvectors. Then a 
sparse linear regression model is fit to each row of the embedding data matrix 
using the remaining rows as predictors, the rows of which are stored in 
matrix $C$. Finally, an affinity matrix $B = |C| + |C|^\top$ is constructed, 
where $|\cdot|$ is applied entry-wise. If the data in the embedding lie 
perfectly in $K$ subspaces, then $B_{ij} = 0$ for $i, j$ not in the same 
subspace. If the embedding of $P$ is used, then this property holds, but it 
is true in general for the embedding of $A$, in which case a final clustering 
step is required. For these simulations, spectral clustering with the 
normalized Laplacian and Gaussian mixture models is used for this final step.

Modularity maximization is NP-hard, so \citet{307cbeb9b1be48299388437423d94bf1} 
used the Extreme Points (EP) algorithm \cite{le2016}, which is $O(n^{K - 1})$. 
For these simulations, the EP algorithm was used for $K = 2$, and for $K > 2$, 
the Louvain algorithm \cite{Blondel_2008} was used instead.

## Balanced communities

In each simulation, community labels $z_1, ..., z_n$ were drawn 
from a multinomial distribution with mixture parameters 
$\{\alpha_1, ..., \alpha_K\}$, then $\{\lambda^{(kl)}\}_K$ according to the 
drawn community labels, $P$ was constructed using the drawn 
$\{\lambda^{(kl)}\}_K$, and $A$ was drawn from $P$ by 
$A_{ij} \stackrel{indep}{\sim} Bernoulli(P_{ij})$. Each simulation has a unique
edge probability matrix $P$.

For these examples, we set the following parameters:

* Number of vertices $n = 128, 256, 512, 1024, 2048, 4096$
* Number of underlying communities $K = 2, 3, 4$
* Mixture parameters $\alpha_k = 1 / K$ for $k = 1, ..., K$, (i.e., each 
community label has an equal probability of being drawn)
* Community labels 
$z_k \iid Multinomial(\alpha_1, ..., \alpha_K)$
* Within-group popularities $\lambda^{(kk)} \iid Beta(2, 1)$
* Between-group popularities 
$\lambda^{(kl)} \iid Beta(1, 2)$ for $k \neq l$

$50$ simulations were performed for each $(n, K)$ pair.

Fig \ref{fig:clust_err_k} show that for large $n$, Algorithm 2 results in 
a misclustering rate of 0. Sparse subspace clustering produces similar results 
for $K > 2$.

```{r clust_err_k, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4. Communities are approximately balanced. Simulations were repeated 50 times for each sample size.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
ssc.df <- readr::read_csv('clustering-ssc-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  dplyr::inner_join(
    ssc.df %>% 
      dplyr::group_by(n, K) %>% 
      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
                       first.q.ssc2 = quantile(error.ssc2, .25),
                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
      dplyr::ungroup()
  ) %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  labs(y = 'community detection error rate', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC-A')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC-A'), width = .1) + 
  geom_line(aes(x = n, y = med.err.mm,
                colour = 'MM-EP')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
                    colour = 'MM-EP'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc2,
                colour = 'SSC-ASE')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
                    colour = 'SSC-ASE'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10()
```

Theorem 4 implies that algorithm 2 will result in not just in the error rate 
converging to 0 but the error *count* as well (Fig. \ref{fig:clust_err_ct_sim}).
We explore this in Fig \ref{fig:clust_err_ct_sim}.

```{r clust_err_ct_sim, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of communities from 2 to 4. Communities are approximately balanced. Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k.csv')
ssc.df <- readr::read_csv('clustering-ssc-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  dplyr::inner_join(
    ssc.df %>% 
      dplyr::group_by(n, K) %>% 
      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
                       first.q.ssc2 = quantile(error.ssc2, .25),
                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
      dplyr::ungroup()
  ) %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  labs(y = 'community detection error count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC-A')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC-A'), width = .1) + 
  geom_line(aes(x = n, y = med.err.mm * n,
                colour = 'MM-EP')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
                    colour = 'MM-EP'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc2 * n,
                colour = 'SSC-ASE')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
                    colour = 'SSC-ASE'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10()
```

Given ground truth community labels, Algorithm 5 and the MLE-based plug-in 
estimators \cite{307cbeb9b1be48299388437423d94bf1} \cite{noroozi2019estimation} 
perform similarly, with root mean square error decaying at rate approximately 
$n^{-1/2}$.

```{r lambda_est_k, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size.', fig.width = 10}
rmse.df <- readr::read_csv('rmse-k.csv')

rmse.df %>% 
  na.omit() %>% 
  dplyr::group_by(K, n) %>% 
  dplyr::summarise(median.rmse = median(rmse),
                   q1.rmse = quantile(rmse, .25),
                   q3.rmse = quantile(rmse, .75),
                   median.rmse.mle = median(rmse.mle),
                   q1.rmse.mle = quantile(rmse.mle, .25),
                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
                    colour = 'Algorithm 4'), width = .1) + 
  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
                   colour = 'MLE-based'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(y = 'RMSE', colour = NULL) + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
```

## Imbalanced communities

Simulations performed in this section are similar to those in the previous 
section with the exception of the mixture parameters 
$\{\alpha_1, ..., \alpha_K\}$ used to draw community labels from the 
multinomial distribution. For these examples, we set the following parameters:

* Number of vertices $n = 128, 256, 512, 1024, 2048, 4096$
* Number of underlying communities $K = 2, 3, 4$
* Mixture parameters $\alpha_k = \frac{k^{-1}}{\sum_{l=1}^K l^{-1}}$ for 
$k = 1, ..., K$
* Community labels 
$z_k \iid Multinomial(\alpha_1, ..., \alpha_K)$
* Within-group popularities $\lambda^{(kk)} \iid Beta(2, 1)$
* Between-group popularities 
$\lambda^{(kl)} \iid Beta(1, 2)$ for $k \neq l$

$50$ simulations were performed for each $(n, K)$ pair.

Fig \ref{fig:clust_err_k_imba} and \ref{fig:clust_err_ct_sim_imba} show 
similar results as in the balanced communities case, with both Algorithm 2 and 
sparse subspace clustering resulting in no misclustered vertices for a 
sufficiently large sample size. However, Fig \ref{fig:lambda_est_k_imba} 
suggests that while Algorithm 5 retains $\sqrt{n}$ efficiency, the MLE-based 
plug-in estimator fails to do so.

```{r clust_err_k_imba, fig.cap = 'IQR of clustering error using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of underlying communities from 2 to 4. Communities are imbalanced. Simulations were repeated 50 times for each sample size.', fig.width = 10}
clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
ssc.df <- readr::read_csv('clustering-ssc-k-imbalanced.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  dplyr::inner_join(
    ssc.df %>% 
      dplyr::group_by(n, K) %>% 
      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
                       first.q.ssc2 = quantile(error.ssc2, .25),
                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
      dplyr::ungroup()
  ) %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  labs(y = 'community detection error rate', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC-A')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC-A'), width = .1) + 
  geom_line(aes(x = n, y = med.err.mm,
                colour = 'MM-EP')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm, ymax = third.q.mm,
                    colour = 'MM-EP'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc2,
                colour = 'SSC-ASE')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
                    colour = 'SSC-ASE'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10()
```

```{r clust_err_ct_sim_imba, fig.cap = 'IQR of counts of misclustered vertices using Algorithm 2 (blue) compared against subspace clustering on the adjacency spectral embedding (purple), the extreme points algorithm (red), and subspace clustering on the adjacency matrix (green), for sample sizes from 128 to 4096 and number of communities from 2 to 4. Communities are imbalanced. Simulations were repeated 50 times for each sample size.', cache = FALSE, fig.width = 10}
clustering.df <- readr::read_csv('clustering-k-imbalanced.csv')
ssc.df <- readr::read_csv('clustering-ssc-k-imbalanced.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(med.err = median(error),
                   first.q = quantile(error, .25),
                   third.q = quantile(error, .75),
                   med.err.ssc = median(error.ssc),
                   first.q.ssc = quantile(error.ssc, .25),
                   third.q.ssc = quantile(error.ssc, .75),
                   med.err.mm = median(error.mm),
                   first.q.mm = quantile(error.mm, .25),
                   third.q.mm = quantile(error.mm, .75)) %>%
  dplyr::ungroup() %>% 
  dplyr::inner_join(
    ssc.df %>% 
      dplyr::group_by(n, K) %>% 
      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
                       first.q.ssc2 = quantile(error.ssc2, .25),
                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
      dplyr::ungroup()
  ) %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  labs(y = 'community detection error count', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC-A')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC-A'), width = .1) + 
  geom_line(aes(x = n, y = med.err.mm * n,
                colour = 'MM-EP')) + 
  geom_errorbar(aes(x = n, ymin = first.q.mm * n, ymax = third.q.mm * n,
                    colour = 'MM-EP'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc2 * n,
                colour = 'SSC-ASE')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc2 * n, ymax = third.q.ssc2 * n,
                    colour = 'SSC-ASE'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10()
```

```{r lambda_est_k_imba, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 (red) compared against an MLE-based method (blue) using sample sizes from 128 to 4096 and number of clusters from 2 to 4. Simulations were repeated 50 times for each sample size.', fig.width = 10}
rmse.df <- readr::read_csv('rmse-k-imbalanced.csv')

rmse.df %>% 
  na.omit() %>% 
  dplyr::group_by(K, n) %>% 
  dplyr::summarise(median.rmse = median(rmse),
                   q1.rmse = quantile(rmse, .25),
                   q3.rmse = quantile(rmse, .75),
                   median.rmse.mle = median(rmse.mle),
                   q1.rmse.mle = quantile(rmse.mle, .25),
                   q3.rmse.mle = quantile(rmse.mle, .75)) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  geom_line(aes(x = n, y = median.rmse, colour = 'Algorithm 4')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse, ymax = q3.rmse,
                    colour = 'Algorithm 4'), width = .1) + 
  geom_line(aes(x = n, y = median.rmse.mle, colour = 'MLE-based')) + 
  geom_errorbar(aes(x = n, ymin = q1.rmse.mle, ymax = q3.rmse.mle,
                   colour = 'MLE-based'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  labs(y = 'RMSE', colour = NULL) + 
  facet_wrap(~ K, labeller = 'label_both') + 
  scale_y_log10() + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096))
```

## Additional experiments

Using the same set of parameters for generating $P$ and $A$ as in the 
balanced communities examples for $K = 2$, we generated one instance of $A$ for
each $n$ and constructed $B$ according to algorithm 4 to verify that 
as $n \to \infty$, $(\hat{v}_i)^\top \hat{v_j} \stackrel{P}{\to} 0$ for $i, j$ 
in different clusters. Furthermore, the distribution of these inner products 
should be approximately normal.

```{r, fig.cap = 'Between-cluster inner products of the eigenvectors of $A$ for varying sample sizes.', fig.height = 2, fig.width = 4}
densities.df <- readr::read_csv('densities.csv')

ggplot(densities.df) + 
  geom_density(aes(x = inner.prods * n, colour = factor(n))) + 
  labs(x = expression(n * v[i]^T * v[j]),
       colour = 'n') + 
  scale_x_continuous(breaks = seq(-5, 5)) + 
  scale_colour_brewer(palette = 'Set1')
```

# Real data examples

```{r}
import::here(cluster.pabm, ssc, plot.A, .from = '~/dev/pabm-grdpg/functions.R')

n.edges <- 20566

butterfly <- R.matlab::readMat('data/Raw_butterfly_network.mat')

A <- butterfly$W.butterfly0
z <- butterfly$labels

keep <- c(6, 2, 9, 4)
A <- A[z %in% keep, z %in% keep]
z <- z[z %in% keep]
n <- length(z)
K <- 4

delta <- quantile(A[upper.tri(A)], 1 - n.edges / (n * (n - 1) / 2))
A <- (A > delta) * 1

z.hat <- cluster.pabm(A, K)

# z.hat.mm <- A %>% 
#   igraph::graph_from_adjacency_matrix(mode = 'undirected') %>% 
#   igraph::cluster_louvain() %>% 
#   igraph::membership()

# ari <- round(fossil::adj.rand.index(z, z.hat) * 100)
```

In the first real data example, we applied Algorithm 2 to the Leeds Butterfly 
dataset \cite{Wang_2018} consisting of visual similarity measurements among 832 
butterflies across 10 species. The graph was modified to match the example 
from \citet{noroozi2019estimation}: Only the 4 most frequent species were 
considered, and the similarities were discretized to $\{0, 1\}$ via 
thresholding. Fig. \ref{fig:butterfly} shows a sorted adjacency matrix sorted 
by the resultant clustering.

Comparing against the ground truth species labels, Algorithm 2 achieves an 
accuracy of 63\% and an adjusted Rand index of 73\%. In comparison, 
\citet{noroozi2019estimation} achieved an adjusted Rand index of 73\% using 
sparse subspace clustering on the same dataset.

```{r butterfly, fig.height = 2, fig.width = 2, fig.cap = 'Adjacency matrix of the Leeds Butterfly dataset after sorting by the clustering outputted by Algorithm 2.', cache = FALSE}
plot.A(A, z.hat)
```

```{r}
# this portion is from the code provided by Sengupta and Chen
# the data were generated by scripts from S&C, then saved as RDS files

bmp <- readRDS('data/bmp.rds')
A.bmp <- bmp$A
z.bmp <- bmp$z

pb <- readRDS('data/pb.rds')
A.pb <- pb$A
z.pb <- pb$z

dblp <- readRDS('data/dblp.rds')
A.dblp <- dblp$A
z.dblp <- dblp$z
```

```{r, cache = TRUE}
zhat.bmp <- cluster.pabm(A.bmp, 2, use.all = TRUE, normalize = FALSE)
bmp.acc <- mean(zhat.bmp == z.bmp)
bmp.acc <- ifelse(bmp.acc < .5, 1 - bmp.acc, bmp.acc)

zhat.pb <- cluster.pabm(A.pb, 2, use.all = TRUE, normalize = FALSE, 
                        p = 2, q = 0, d.eigenmap = 2)
pb.acc <- mean(zhat.pb == z.pb)
pb.acc <- ifelse(pb.acc < .5, 1 - pb.acc, pb.acc)

zhat.dblp <- cluster.pabm(A.dblp, 2, use.all = TRUE, normalize = FALSE)
dblp.acc <- mean(zhat.dblp == z.dblp)
dblp.acc <- ifelse(dblp.acc < .5, 1 - dblp.acc, dblp.acc)

zhat.bmp.ssc <- ssc(A.bmp, 2, 1e-3, normalize = TRUE)
bmp.acc.ssc <- mean(zhat.bmp.ssc == z.bmp)
bmp.acc.ssc <- ifelse(bmp.acc.ssc < .5, 1 - bmp.acc.ssc, bmp.acc.ssc)

zhat.pb.ssc <- ssc(A.pb, 2, .008, normalize = TRUE)
pb.acc.ssc <- mean(zhat.pb.ssc == z.pb)
pb.acc.ssc <- ifelse(pb.acc.ssc < .5, 1 - pb.acc.ssc, pb.acc.ssc)

zhat.dblp.ssc <- ssc(A.dblp, 2, 1e-2, normalize = FALSE)
dblp.acc.ssc <- mean(zhat.dblp.ssc == z.dblp)
dblp.acc.ssc <- ifelse(dblp.acc.ssc < .5, 1 - dblp.acc.ssc, dblp.acc.ssc)

```

In the second example, we applied Algorithm 2 to the British MPs Twitter 
network \cite{greene2013producing}, the Political Blogs network 
\cite{10.1145/1134271.1134277}, and the DBLP network \cite{NIPS2009_3855} 
\cite{10.1007/978-3-642-15880-3_42}. For this data analysis, we subsetted the 
data as described by 
\citet{307cbeb9b1be48299388437423d94bf1} for their analysis of the same 
networks. Our methods underperformed compared to modularity maximization, 
although performance is comparable. In 
addition, Algorithm 2's runtime is much lower than that of modularity 
maximization.

```{r}
dplyr::tibble(Network = c('British MPs', 'Political blogs', 'DBLP'),
              Vertices = c(nrow(A.bmp), nrow(A.pb), nrow(A.dblp)),
              `Error (Mod. Max.)` = c(round(1 / 329, 3),
                                      round(61 / 1222, 3),
                                      round(62 / 2203, 3)),
              `Error (SSC)` = round(1 - c(bmp.acc.ssc,
                                          pb.acc.ssc,
                                          dblp.acc.ssc), 3),
              `Error (Alg. 2)` = round(1 - c(bmp.acc, 
                                                  pb.acc,
                                                  dblp.acc), 3)) %>% 
  knitr::kable(caption = paste('Community detection error rates', 
                               'for modularity maximization,',
                               'sparse subspace clustering, ',
                               'and Alg. 2.'))
```

```{r mp, cache = FALSE, fig.width = 8, fig.height = 2, fig.cap = 'Adjacency matrices of (from left to right) the British MPs, Political Blogs, and DBLP networks after sorting by the clustering outputted by Algorithm 2.', fig.dpi = 10}
gridExtra::grid.arrange(plot.A(A.bmp, zhat.bmp),
                        plot.A(A.pb, zhat.pb),
                        plot.A(A.dblp, zhat.dblp),
                        ncol = 3)
```

In the third example, we consider the Karantaka villages data studied by 
\citet{DVN/U3BIHX_2013}. For this example, we chose the `visitgo` networks from 
villages 12, 31, and 46 at the household level. The label of interest is the 
religious affiliation. The networks were truncated to religions "1" and "2", 
and vertices of degree 0 were removed.

```{r, cache = FALSE, fig.cap = 'Adjacency matrix of the Karnataka villages data, arranged by the clustering produced by Algorithm 2 (left). The villages studied here are, from left to right, 12, 31, and 46.', fig.height = 2, fig.width = 8, eval = TRUE}
K <- 2
hh.12 <- readRDS('data/village-hh-12.rds')
hh.31 <- readRDS('data/village-hh-31.rds')
hh.46 <- readRDS('data/village-hh-46.rds')

# this part of code taken from sengupta & chen
setwd('~/dev/pabm-grdpg/codes_and_data')
require(MASS);require("clue")
source("EPalgosim.R") 
source("functions.R")
mm <- function(A.hh) {
  b.can = EPalgo(A.hh,eps=0) # EP algorithm (no perturbation)
  Q.PA.can = rep(NA, ncol(b.can))	# array to store Q values
  for (i in 1:ncol(b.can)){
    #check if any cluster is empty
    foo = rep(NA, 2)
    for (clus in 1:2) {foo[clus]=sum(b.can[,i]==clus)}
    if (min(foo)==0) {stop('Empty groups are not allowed')} 
    Q.PA.can[i] = Q.PA(A.hh, b=b.can[,i])   # fit PABM
  } # end of i for loop
  foo1 = order(-Q.PA.can)[1] 
  return(b.can[,foo1])   # community assignment that maximises Q.PA
}

A.hh.12 <- hh.12$A
z.hh.12 <- hh.12$z
zhat.hh.12 <- cluster.pabm(A.hh.12, K,
                           use.all = TRUE, normalize = FALSE, p = 3, q = 1)
zhat.hh.ssc.12 <- ssc(A.hh.12, K, lambda = .1)
zhat.hh.mm.12 <- mm(A.hh.12)

A.hh.31 <- hh.31$A
z.hh.31 <- hh.31$z
zhat.hh.31 <- cluster.pabm(A.hh.31, K,
                           use.all = TRUE, normalize = FALSE, p = 3, q = 1)
zhat.hh.ssc.31 <- ssc(A.hh.31, K, lambda = .05)
zhat.hh.mm.31 <- mm(A.hh.31)

A.hh.46 <- hh.46$A
z.hh.46 <- hh.46$z
zhat.hh.46 <- cluster.pabm(A.hh.46, K,
                           use.all = TRUE, normalize = FALSE, p = 2, q = 0)
zhat.hh.ssc.46 <- ssc(A.hh.46, K, lambda = .0005)
zhat.hh.mm.46 <- mm(A.hh.46)

gridExtra::grid.arrange(plot.A(A.hh.12, zhat.hh.12),
                        plot.A(A.hh.31, zhat.hh.31),
                        plot.A(A.hh.46, zhat.hh.46),
                        nrow = 1)
```

```{r, eval = TRUE}
hh.acc.12 <- mean(zhat.hh.12 == z.hh.12)
hh.acc.12 <- ifelse(hh.acc.12 < .5, 1 - hh.acc.12, hh.acc.12)
hh.acc.ssc.12 <- mean(zhat.hh.ssc.12 == z.hh.12)
hh.acc.ssc.12 <- ifelse(hh.acc.ssc.12 < .5, 1 - hh.acc.ssc.12, hh.acc.ssc.12)
hh.acc.mm.12 <- mean(zhat.hh.mm.12 == z.hh.12)
hh.acc.mm.12 <- ifelse(hh.acc.mm.12 < .5, 1 - hh.acc.mm.12, hh.acc.mm.12)

hh.acc.31 <- mean(zhat.hh.31 == z.hh.31)
hh.acc.31 <- ifelse(hh.acc.31 < .5, 1 - hh.acc.31, hh.acc.31)
hh.acc.ssc.31 <- mean(zhat.hh.ssc.31 == z.hh.31)
hh.acc.ssc.31 <- ifelse(hh.acc.ssc.31 < .5, 1 - hh.acc.ssc.31, hh.acc.ssc.31)
hh.acc.mm.31 <- mean(zhat.hh.mm.31 == z.hh.31)
hh.acc.mm.31 <- ifelse(hh.acc.mm.31 < .5, 1 - hh.acc.mm.31, hh.acc.mm.31)

hh.acc.46 <- mean(zhat.hh.46 == z.hh.46)
hh.acc.46 <- ifelse(hh.acc.46 < .5, 1 - hh.acc.46, hh.acc.46)
hh.acc.ssc.46 <- mean(zhat.hh.ssc.46 == z.hh.46)
hh.acc.ssc.46 <- ifelse(hh.acc.ssc.46 < .5, 1 - hh.acc.ssc.46, hh.acc.ssc.46)
hh.acc.mm.46 <- mean(zhat.hh.mm.46 == z.hh.46)
hh.acc.mm.46 <- ifelse(hh.acc.mm.46 < .5, 1 - hh.acc.mm.46, hh.acc.mm.46)

knitr::kable(dplyr::tibble(
  Method = c('Alg. 2', 'SSC', 'Mod. Max.'),
  `Error (Village 12)` = round(1 - c(hh.acc.12, hh.acc.ssc.12, hh.acc.mm.12), 3),
  `Error (Village 31)` = round(1 - c(hh.acc.31, hh.acc.ssc.31, hh.acc.mm.31), 3),
  `Error (Village 46)` = round(1 - c(hh.acc.46, hh.acc.ssc.46, hh.acc.mm.46), 3)),
  caption = paste('Community detection error rates for identifying',
                  'household religion.'))
```

# Discussion

# References