### 2021-04-03 short-summary.Rmd
wrote a short summary of work thus far + future work

---
title: 'Summary of Dissertation Work Thus Far and Future Steps'
author: John Koo
output:
  pdf_document:
    citation_package: natbib
# output: html_document
urlcolor: blue
header-includes:
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square}
- \usepackage{verbatim}
- \usepackage{amsthm}
bibliography: short-summary.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.pos = 'H', 
                      fig.align = 'center', 
                      fig.height = 5, 
                      fig.width = 12, 
                      fig.dpi = 300)
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
```

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{indep}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\theoremstyle{example}
\newtheorem*{example}{Example}

# Introduction and Research Goal/Scope

Statistical analysis on graphs or networks often involves the partitioning of a graph into disconnected subgraphs or clusters, i.e., finding some function $f : [n] \mapsto [K]$. This is often motivated by the assumption that there exist underlying and unobserved communities to which each vertex of the graph belongs, and edges between pairs of vertices are determined by drawing from a probability distribution based on the community relationships between each pair. The goal of the partitioning or clustering then is community detection, or the recovery of the true underlying community labels for each vertex, up to permutation (with some additional parameter estimation being of possible interest), assuming some underlying probability model.

One example of such a model is the Stochastic Block Model (SBM) with two communities. In this case, we may draw edges between communities independently such that the distribution of each edge within community 1 is $Bernoulli(p)$, the distribution of each edge within community 2 is $Bernoulli(q)$, and the distribution of each edge between communities 1 and 2 is $Bernoulli(r)$. Other models have been proposed, such as the Degree-Corrected Block Model (DCBM), Mixed Membership Stochastic Block Model (MMSBM), and Popularity Adjusted Block Model (PABM), which are often considered generalizations of the SBM. All such models involve undirected, unweighted, hollow graphs (although there are some straightforward generalizations to other types of graphs, in particular to weighted graphs). The information in this graph is stored in adjacency $A \in \{0, 1\}^{n \times n}$ where $A_{ij} = 1$ if there is an edge between vertices $i$ and $j$, and 0 otherwise.

The underlying similarity among all of these Bernoulli edge graphs is the edge probability matrix, $P \in [0, 1]^{n \times n}$. Each element $P_{ij}$ is the probability of the existence of an edge between vertices $i$ and $j$ ($P$ is then symmetric and $P_{ii} = 0$ $\forall i$). The adjacency matrix $A$ is then drawn such that $A_{ij} \stackrel{indep}{\sim} P_{ij}$ for each $i < j$ and $A_{ji} = A_{ij}$. For instance, in the SBM example from before, $P_{ij} = p$ if vertices $i$ and $j$ both belong to community 1, $P_{ij} = q$ if they both belong to community 2, and $P_{ij} = r$ if one is in community 1 and the other is in community 2.

## The Popularity Adjusted Block Model

\begin{definition}[Popularity Adjusted Block Model]
\label{pabm}
Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
set of $n$ 
vertices, $V$. Each vertex has a community label $1, ..., K$, and the rows and 
columns of $P$ are arranged by community label such that $n_k \times n_l$ block 
$P^{(kl)}$ describes the edge probabilities between vertices in communities 
$k$ and $l$ ($P^{(lk)} = (P^{(kl)})^\top$). 
Let graph $G = (V, E)$ be an undirected, unweighted graph such 
that its corresponding adjacency matrix $A \in \{0, 1\}^{n \times n}$ is a 
realization of $\Bernoulli(P)$, i.e., 
$A_{ij} \indep \Bernoulli(P_{ij})$ for $i > j$ 
($A_{ij} = A_{ji}$ and $A_{ii} = 0$).

If each block $P^{(kl)}$ can be written as the outer product of two vectors:

\begin{equation} \label{eq:pabm}
  P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top
\end{equation}

for a set of $K^2$ fixed vectors $\{\lambda^{(st)}\}_{s, t = 1}^K$ where each 
$\lambda^{(st)}$ is a column vector 
of dimension $n_s$, then graph $G$ and its corresponding adjacency matrix $A$ 
is a realization of a popularity adjusted block model with parameters 
$\{\lambda^{(st)}\}_{s, t = 1}^K$. 
\end{definition}

We will use the notation $A \sim PABM(\{\lambda^{(kl)}\}_K)$ to denote a random 
adjacency matrix $A$ drawn from a PABM with parameters $\lambda^{(kl)}$ 
consisting of $K$ underlying communities.

## The Generalized Random Dot Product Graph

The Random Dot Product Graph (RDPG) is another graph model with Bernoulli edge probabilities. Under this model, each vertex of the graph can be represented by a point in some latent space such that the edge probability between any pair of vertices is given by their corresponding dot product in the latent space. It can be shown that any Bernoulli edge graph with positive semidefinite edge probability matrix $P$ can be represented as a RDPG. More specifically, if $P = V \Sigma V^\top$ is the spectral decomposition of $P$ such that $\Sigma$ is a diagonal matrix of $d$ positive eigenvalues, then $V \Sigma^{1/2}$ is one representation of the latent positions under the RDPG framework.

\begin{example}[Connecting the SBM to the RDPG]
Consider the same SBM as before. Let community 1 have $n_1$ vertices and community 2 have $n_2$ vertices such that $n_1 + n_2 = n$. Let the edge probability matrix $P$ be organized by community such that the $P^{(kl)}$ block represents edge probabilities between communities $k$ and $l$. In this case, $P = \begin{bmatrix} P^{(11)} & P^{(12)} \\ P^{(21)} & P^{(22)} \end{bmatrix}$ where each block is a constant value, e.g., $P^{(11)}_{ij} = p$. Then one RDPG representation of this SBM is:

$$X = \begin{bmatrix} 
\sqrt{p} & 0 \\
\vdots & \vdots \\
\sqrt{p} & 0 \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\ 
\vdots & \vdots \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
\end{bmatrix}
\in \mathbb{R}^{n \times 2}$$

where the first $n_1$ rows are $\begin{bmatrix} \sqrt{p} & 0 \end{bmatrix}$ and the next $n_2$ rows are $\begin{bmatrix} \sqrt{r^2 / p} & \sqrt{q - r^2 / p} \end{bmatrix}$. Then it can be shown that

$$P = X X^\top$$
\end{example}

If $P$ is not positive semidefinite, then it cannot be the edge probability 
matrix of a RDPG, but it can be the edge probability matrix of a *Generalized*
Random Dot Product Graph (GRDPG).[^grdpg]

[^grdpg]: It is trivial to show that *all* Bernoulli graphs are GRDPGs.

\begin{definition}[Generalized Random Dot Product Graph] 
\label{grdpg} 
Let $P \in [0, 1]^{n \times n}$ be a symmetric edge probability matrix for a 
set of $n$ vertices, $V$. If $\exists X \in \mathbb{R}^{n \times d}$ such that

\begin{equation} \label{eq:grdpg}
  P = X I_{pq} X^\top
\end{equation}

for some $d, p, q \in \mathbb{N}$ and $p + q = d$, then 
graph $G = (V, E)$ with adjacency matrix $A$ such that 
$A_{ij} \indep \Bernoulli(P_{ij})$ for $i > j$ ($A_{ij} = A_{ji}$ and 
$A_{ii} = 0$) is a draw from the generalized random dot product graph model 
with latent positions $X$ and signature $(p, q)$. More precisely, if vertices 
$i$ and $j$ have latent positions $x_i$ and $x_j$ respectively, then the edge 
probability between the two is $P_{ij} = x_i^\top I_{pq} x_j$, and $X$ contains 
the latent positions as rows $x_i^\top$. 
\end{definition}

We will use the notation $A \sim GRDPG_{p,q}(X)$ to denote a random adjacency 
matrix $A$ drawn from latent positions $X$ and signature $(p, q)$.

\begin{definition}[Indefinite Orthogonal Group] 
The indefinite orthogonal group with signature $(p, q)$ is 
the set $\{Q \in \mathbb{R}^{d \times d} : Q I_{pq} Q^{\top} = I_{pq}\}$, 
denoted as $\mathbb{O}(p, q)$.
\end{definition}

\begin{remark}
Like the RDPG, the latent positions of a GRDPG are not unique. 
More specifically, if $P_{ij} = x_i^\top I_{pq} x_j$, then we also have for any 
$Q \in \mathbb{O}(p, q)$, 
$(Q x_i)^\top I_{pq} (Q x_j) = x_i^\top (Q^\top I_{pq} Q) x_j = 
x_i^\top I_{pq} x_j = P_{ij}$. 
Unlike in the RDPG case, transforming the latent positions via multiplication 
by $Q \in \mathbb{O}(p, q)$ does not necessarily maintain interpoint angles or 
distances. 
\end{remark}

# Literature Review and Previous Work

The PABM was introduced by \citet{307cbeb9b1be48299388437423d94bf1}, who 
intuited that a spectral clustering type of approach would be appropriate for 
this problem. Their work was furthered by \citet{noroozi2019estimation}, who 
used SSC for community detection. SSC is typically performed on points in 
Euclidean space, which suggests a spectral embedding, but the authors of this 
paper instead performed SSC on $P$ and $A$ directly. They were able to show 
that SSC on $P$ results in perfect community detection.

SSC is described as follows. Given $x_1, ..., x_n \in \mathbb{R}^d$ such 
that each point lies on one of $K$ subspaces, the goal is to determine 
which subspace each point belongs to. We might also want to add some noise to 
the points so that they don't lie exactly on their respective subspaces.

SSC is performed by solving an optimization problem for each observed point.
Given $X \in \mathbb{R}^{n \times d}$ with vectors 
$x_i^\top \in \mathbb{R}^d$ as rows of $X$, the optimization problem 
$\min_{c_i} ||c_i||_1$ subject to $x_i = X c_i$ and $\beta_i = 0$ is solved 
for each $i = 1, ..., n$. The solutions are collected collected into matrix 
$C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}^\top$ to construct an 
affinity matrix $B = |C| + |C^\top|$. If each $x_i$ lie perfectly on one of $K$ 
subspaces, $B$ describes an undirected graph consisting of $K$ disjoint 
subgraphs, i.e., $B_{ij} = 0$ if $x_i, x_j$ are in different subspaces. 
If $X$ instead represents points near $K$ subspaces with some noise, a final 
graph partitioning step is performed (e.g., edge thresholding or spectral 
clustering).

In practice, SSC is often performed by solving the LASSO problems

\begin{equation} \label{eq:ssc}
c_i = \arg\min_c \frac{1}{2} ||x_i - X_{-i} c||^2_2 + \lambda ||c||_1
\end{equation}

for some sparsity parameter $\lambda > 0$. The $c_i$ vectors are then collected 
into $C$ and $B$ as before.

\begin{definition}[Subspace Detection Property] 
Let $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$ be noisy 
points sampled from $K$ subspaces. Let $C$ and $B$ be constructed from the 
solutions of LASSO problems as described in (\ref{eq:ssc}). If each column of 
$C$ has nonzero norm and $B_{ij} = 0$ $\forall$ $x_i$ and $x_j$ sampled from 
different subspaces, then $X$ obeys the subspace detection property. 
\end{definition}

\citet{pmlr-v28-wang13} showed that in the case where the subspaces are 
orthogonal (or at least the cosine of their angles is very small) and the 
points are close enough to their respective subspaces, the subspace detection
property will hold given an appropriate choice of $\lambda$. It turns out that 
this is precisely the case for a particular embedding choice for the PABM, of 
which the properties have been described by 
\citet{rubindelanchy2017statistical}.

# Completed Work Thus Far

The PABM can be expressed as a GRDPG. In particular, there is one expression 
of this that results in a very simple structure such that each community 
corresponds to a subspace and the subspaces are mutually orthogonal.

\begin{theorem}[Connecting the PABM to the GRDPG for $K = 2$]
\label{theorem1}  
Let

$$X = \begin{bmatrix} 
\lambda^{(11)} & \lambda^{(12)} & 0 & 0 \\ 
0 & 0 & \lambda^{(21)} & \lambda^{(22)} 
\end{bmatrix}$$

$$U = \begin{bmatrix} 1 & 0 & 0 & 0 \\
0 & 0 & 1 / \sqrt{2} & 1 / \sqrt{2} \\
0 & 0 & 1 / \sqrt{2} & - 1 / \sqrt{2} \\
0 & 1 & 0 & 0 \end{bmatrix}$$

where each $\lambda^{(kl)}$ is a vector as in Definition 1. 
Then $A \sim GRDPG_{3, 1}(X U)$ and $A \sim PABM(\{(\lambda^{(kl)}\}_2)$ are 
equivalent.
\end{theorem}

\begin{theorem}[Generalization to $K > 2$] 
\label{theorem2}
There exists a block diagonal matrix 
$X \in \mathbb{R}^{n \times K^2}$ defined by PABM parameters 
$\{\lambda^{(kl)}\}_K$ and orthonormal matrix 
$U \in \mathbb{R}^{K^2 \times K^2}$ that is fixed 
for each $K$ such that $A \sim GRDPG_{K (K+1) / 2, K (K-1) / 2}(XU)$ and 
$A \sim PABM(\{(\lambda^{(kl)}\})_K)$ are equivalent.
\end{theorem}

\begin{proof}
Define the following matrices from $\{\lambda^{(kl)}\}_K$:

$$\Lambda^{(k)} = 
\begin{bmatrix} \lambda^{(k,1)} & \cdots & \lambda^{(k, K)} \end{bmatrix}
\in \mathbb{R}^{n_k \times K}$$

\begin{equation} \label{eq:xy}
X = \blockdiag(\Lambda^{(1)}, ..., \Lambda^{(K)}) \in \mathbb{R}^{n \times K^2}
\end{equation}

$$L^{(k)} = \blockdiag(\lambda^{(1k)}, ..., \lambda^{(Kk)}) \in 
\mathbb{R}^{n \times K}$$

$$Y = \begin{bmatrix} L^{(1)} & \cdots & L^{(K)} \end{bmatrix} \in 
\mathbb{R}^{n \times K^2}$$

Then $P = X Y^\top$.

Similar to the $K = 2$ case, we have $Y = X \Pi$ for a permutation matrix
$\Pi$, resulting in $P = X \Pi X^\top$.  
The permutation described by $\Pi$ has $K$ fixed points, which correspond to 
$K$ eigenvalues equal to $1$ with corresponding eigenvectors $e_k$ where 
$k = r (K + 1) + 1$ for $r = 0, ..., K - 1$. It also has 
$\binom{K}{2} = K (K - 1) / 2$ cycles of order $2$. Each cycle corresponds to 
a pair of eigenvalues $+1$ and $-1$ and a pair of eigenvectors 
$(e_s + e_t) / \sqrt{2}$ and $(e_s - e_t) / \sqrt{2}$.

Then $\Pi$ has $K (K + 1) / 2$ eigenvalues equal to $1$ and $K (K - 1) / 2$ 
eigenvalues equal to $-1$. $\Pi$ has the decomposed form

\begin{equation} \label{eq:permutation}
\Pi = U I_{K (K + 1) / 2, K (K - 1) / 2} U^\top
\end{equation}

The edge probability matrix then can be written as:

\begin{equation} \label{eq:pabm-grdpg}
P = X U I_{p, q} (X U)^\top
\end{equation}

\begin{equation} \label{eq:p}
p = K (K + 1) / 2
\end{equation}

\begin{equation} \label{eq:q}
q = K (K - 1) / 2
\end{equation}

and we can describe the PABM with $K$ communities as a GRDPG with latent 
positions $X U$ with signature $\big( K (K + 1) / 2, K (K - 1) / 2 \big)$.
\end{proof}

\begin{example}[$K = 3$] Using the same notation as in Theorem \ref{theorem2}:

$$X = \begin{bmatrix} 
\lambda^{(11)} & \lambda^{(12)} & \lambda^{(13)} & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & \lambda^{(21)} & \lambda^{(22)} & \lambda^{(23)} & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & \lambda^{(31)} & \lambda^{(32)} & \lambda^{(33)}
\end{bmatrix}$$

$$Y = \begin{bmatrix} 
\lambda^{(11)} & 0 & 0 & \lambda^{(12)} & 0 & 0 & \lambda^{(13)} & 0 & 0 \\
0 & \lambda^{(21)} & 0 & 0 & \lambda^{(22)} & 0 & 0 & \lambda^{(23)} & 0 \\
0 & 0 & \lambda^{(31)} & 0 & 0 & \lambda^{(32)} & 0 & 0 & \lambda^{(33)}
\end{bmatrix}$$

Then $P = X Y^\top$ and $Y = X \Pi$ where $\Pi$ is a permutation matrix 
consisting of $3$ fixed points and $3$ cycles of order 2:

$$\Pi = \begin{bmatrix} 
1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1
\end{bmatrix}$$

* Positions 1, 5, 9 are fixed.

* The cycles of order 2 are $(2, 4)$, $(3, 7)$, and $(6, 8)$.

Therefore, we can decompose $\Pi = U I_{6, 3} U^\top$ where the first three 
columns of $U$ consist of $e_1$, $e_5$, and $e_9$ corresponding to the fixed 
positions $1$, $5$, and $9$, the next three columns consist of eigenvectors 
$(e_k + e_l) / \sqrt{2}$, and the last three columns consist of eigenvectors 
$(e_k - e_l) / \sqrt{2}$, where pairs $(k, l)$ correspond to the cycles of 
order 2 described above.

The latent positions are the rows of  
$$XU = \begin{bmatrix}
  \lambda^{(11)} & 0 & 0 & 
  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 & 
  \lambda^{(12)} / \sqrt{2} & \lambda^{(13)} / \sqrt{2} & 0 \\
  0 & \lambda^{(22)} & 0 & 
  \lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} & 
  -\lambda^{(21)} / \sqrt{2} & 0 & \lambda^{(23)} / \sqrt{2} \\
  0 & 0 & \lambda^{(33)} & 
  0 & \lambda^{(31)} / \sqrt{2} & \lambda^{(32)} / \sqrt{2} & 
  0 & -\lambda^{(31)} / \sqrt{2} & -\lambda^{(32)} / \sqrt{2}
\end{bmatrix}$$
\end{example}

This connection leads to the following property:

\begin{theorem}
\label{theorem3}
Let $P = V D V^\top$ be the spectral decomposition of the edge probability
matrix. Let $B = n V V^\top$. Then $B_{ij} = 0$ if vertices $i$ and $j$ are 
from different communities.
\end{theorem}

In practice, however, we observe $A$ rather than $P$. Using the asymptotic
properties of the Adjacency Spectral Embedding (ASE) of the GRDPG
\cite{rubindelanchy2017statistical}, it can be shown that constructing 
$\hat{B}_n$ from the eigenvectors of $A$ will result in entries of $\hat{B}$
corresponding to between-community pairings going to 0 with probability 1.

\begin{algorithm}[t]
  \DontPrintSemicolon
  \SetAlgoLined
  \KwData{Adjacency matrix $A$, number of communities $K$}
  \KwResult{Community assignments $1, ..., K$}
    Compute the eigenvectors of $A$ that correspond to the $K (K+1) / 2$ most 
    positive eigenvalues and $K (K-1) / 2$ most negative eigenvalues. Construct 
    $V$ using these eigenvectors as its columns.\;
    Compute $B = |n V V^\top|$, applying $|\cdot|$ entry-wise.\;
    Construct graph $G$ using $B$ as its similarity matrix.\;
    Partition $G$ into $K$ disconnected subgraphs  
    (e.g., using edge thresholding or spectral clustering).\;
    Map each partition to the community labels $1, ..., K$.\;
  \caption{Orthogonal Spectral Clustering.}
\end{algorithm}

\begin{theorem}
\label{theorem4} 
Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC 
(Alg. 1). Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\{(\log n)^{4c}\}$,

\begin{equation} \label{eq:thm4}
\max_{i, j} |n (\hat{v}_n^{(i)})^\top \hat{v}_n^{(j)}| = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
\end{equation}

This provides the result that for $i, j$ in different communities, 
$\hat{B}_n^{(ij)} \stackrel{a.s.}{\to} 0$.
\end{theorem}

Theorems \ref{theorem2}, \ref{theorem3}, and \ref{theorem4} 
also provide a very natural path toward using SSC for 
community detection for the PABM. We established in Theorem \ref{theorem2}
that an ASE of the edge probability matrix $P$ can be constructed 
such that the communities lie on mutually orthogonal subspaces,
and this property can be recovered from the eigenvectors of $P$. 
Then Theorems \ref{theorem3} and \ref{theorem4} show that this property holds 
for the unscaled ASE of $A$ drawn from $P$ as $n \to \infty$.

\begin{algorithm}[t]
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{Sparse Subspace Clustering using LASSO \cite{pmlr-v28-wang13}.}
  \KwData{Adjacency matrix $A$, number of communities $K$, 
  hyperparameter $\lambda$}
  \KwResult{Community assignments $1, ..., K$}
    Find $V$, the matrix of eigenvectors of $A$ 
    corresponding to the $K (K + 1) / 2$ most positive 
    and the $K (K - 1) / 2$ most negative eigenvalues.\;
    \For {$i = 1, ..., n$} {
      Assign $v_i^\top$ as the $i^{th}$ row of $V$. 
      Assign $V_{-i} = \begin{bmatrix} 
      v_1 & \cdots & v_{i-1} & v_{i+1} & \cdots & v_n \end{bmatrix}^\top$.\;
      Solve the LASSO problem 
      $c_i = \arg\min_{\beta} 
      \frac{1}{2} ||\sqrt{n} v_i - \sqrt{n} V_{-i} \beta||_2^2 + 
      \lambda ||\beta||_1$.\;
      Assign $\tilde{c}_i = \begin{bmatrix} 
      c_i^{(1)} & \cdots & c_i^{(i-1)} & 0 & c_i^{(i)} & \cdots & c_i^{(n-1)}
      \end{bmatrix}^\top$ such that the superscript is the index of 
      $\tilde{c}_i$.\;
    }
    Assign 
    $C = \begin{bmatrix} \tilde{c}_1 & \cdots & \tilde{c}_n \end{bmatrix}$.\;
    Compute the affinity matrix $B = |C| + |C^\top|$.\;
    Construct graph $G$ using $B$ as its similarity matrix.\;
    Partition $G$ into $K$ disconnected subgraphs (e.g., using edge 
    thresholding or spectral clustering).\;
    Map each partition to the community labels $1, ..., K$. 
\end{algorithm}

\begin{theorem}
\label{theorem5}
Let $P_n$ describe the edge probability matrix of the PABM with 
$n$ vertices, and let $A_n \sim \Bernoulli(P_n)$.  Let $\hat{V}_n$ be the 
matrix of eigenvectors of $A_n$ corresponding to the $K (K + 1) / 2$ most
positive and $K (K - 1) / 2$ most negative eigenvalues. Then 
$\exists \lambda > 0$ such that as $n \to \infty$, $\hat{V}_n$ obeys the 
subspace detection property with probability 1.
\end{theorem}

We are also able to show convergence properties for parameter estimation using 
the ASE, the details of which have been omitted for brevity but can be found 
in the main paper.

# Future Work

So far, we've shown that there is a very convenient GRDPG representation of 
the PABM that makes community detection almost trivially simple. 
More specifically, we showed that an embedding of the PABM results in 
orthogonal subspaces and noise that goes to 0 with probability 1, which makes 
for a very straightforward application of SSC. It can also be shown that this 
type of embedding has the same properties of the SBM and DCBM, which are really
just special cases of the PABM.

Furthermore, all Bernoulli graphs can be
represented as a GRDPG, and it may be the case that other, more complicated 
graphs have nice ASE representations that lead to community detection methods. 
In particular, we are currently exploring the nature of RDPGs and GRDPGs that
result from latent positions for which each community lies on a manifold rather
than on a subspace. 
Much of this work is based on work by \citet{trosset2020learning} who showed 
convergence properties for graphs that result from latent positions that lie 
on a one-dimensional manifold.

\newpage

# References