### 2021-06-09 sdss-talk/slides-4min.Rmd
maybe modify parameter estimation stuff

---
title: Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph
subtitle: SDSS 2021 Lightning Presentation
author: |
  | John Koo, Indiana University
  | Minh Tang, NC State University
  | Michael Trosset, Indiana University
date: 'June 2021'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \setbeamertemplate{itemize items}[circle]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

source('~/dev/pabm-grdpg/functions.R')
set.seed(314159)
```

## The Popularity Adjusted Block Model

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}

```{r out.width = '50%', fig.height = 3, fig.width = 4}
Pz <- generate.P.beta(64, 4, 4, 1, 1, 4)
P <- Pz$P
z <- Pz$clustering
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
```

**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$.  
Then $A \sim \BG(P)$ is a PABM if each 
$P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$

## The Popularity Adjusted Block Model

**Lemma** (Noroozi, Rimal, and Pensky, 2020):

$A$ is sampled from a PABM if $P$ can be described as:

1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
between communities $k$ and $l$. 
2. Organize popularity parameters as vectors 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
of the $i$^th^ vertex of community $k$ towards community $l$. 
3. Each block can be decomposed as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.

**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.

## The Generalized Random Dot Product Graph

Generalized Random Dot Product Graph $A \sim \GRDPG_{p, q}(X)$  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = \blockdiag(I_p, -I_q)$
* $A \sim \BG(X I_{p, q} X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

If latent vectors $X_1, ..., X_n \stackrel{iid}{\sim} F$, then we write 
$(A, X) \sim \GRDPG_{p, q}(F, n)$.

## Connecting the PABM to the GRDPG

**Theorem** (KTT): $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim \GRDPG_{p, q}(X U)$ such that

* $p = K (K + 1) / 2$
* $q = K (K - 1) / 2$
* $U$ is orthogonal and predetermined for each $K$
* $X$ is block diagonal and composed of $\{\lambda^{(kl)}\}_K$  
$\implies$ if $x_i^\top$ and $x_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $x_i^\top x_j = 0$.

**Remark** (non-uniqueness of the latent configuration):  
$A \sim \GRDPG_{p, q}(X U) \implies A \sim \GRDPG_{p, q}(X U Q)$ 
$\forall Q \in \mathbb{O}(p, q)$

**Corollary**: $X$ is block diagonal by community and 
$U$ is orthogonal $\implies$ 
each community corresponds to a subspace in $\mathbb{R}^{K^2}$.

Subspace property holds even with linear transformation 
$Q \in \mathbb{O}(p, q)$.

## Connecting the PABM ot the GRDPG

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ is the spectral decomposition of $P$ for the PABM and 
$V$ has rows $v_i^\top$, then $v_i^\top v_j = 0$ $\forall z_i \neq z_j$.

**Theorem** (KTT): 
If $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ is the spectral decomposition 
of $A$ for the PABM using the $K (K+1) / 2$ most positive and $K (K-1) / 2$ 
most negative eigenvalues, and we denote $\hat{v}_i^\top$ as the rows of 
$\hat{V}$, then

$$\max_{i, j : z_i \neq z_j} n \hat{v}_i^\top \hat{v}_j = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$$

Orthogonal Spectral Clustering (KTT):

1. Let $V$ be the eigenvectors of $A$ corresponding to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $\hat{G}$ using $B$ as its similarity matrix.
4. Partition $\hat{G}$ into $K$ disconnected subgraphs.

## Simulation Study

Simulation setup:

1. $Z_1, ..., Z_n \stackrel{iid}{\sim} \Cat(1 / K, ..., 1 / K)$
2. $\lambda_{ik} \stackrel{iid}{\sim} \Betadist(a_{ik}, b_{ik})$  
$a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$, 
$b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
4. $A \sim \BG(P)$

```{r clust_err_k, fig.width = 10, fig.height = 3, out.width = '100%'}
setwd('..')
clustering.df <- readr::read_csv('clustering-k.csv')
ssc.df <- readr::read_csv('clustering-ssc-k.csv')
clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error),
    first.q = quantile(error, .25),
    third.q = quantile(error, .75),
    med.err.ssc = median(error.ssc),
    first.q.ssc = quantile(error.ssc, .25),
    third.q.ssc = quantile(error.ssc, .75),
    med.err.ep = median(error.ep, na.rm = TRUE),
    first.q.ep = quantile(error.ep, .25, na.rm = TRUE),
    third.q.ep = quantile(error.ep, .75, na.rm = TRUE),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  dplyr::inner_join(
    ssc.df %>% 
      dplyr::group_by(n, K) %>% 
      dplyr::summarise(med.err.ssc2 = median(error.ssc2),
                       first.q.ssc2 = quantile(error.ssc2, .25),
                       third.q.ssc2 = quantile(error.ssc2, .75)) %>% 
      dplyr::ungroup()
  ) %>% 
  ggplot() +
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096)) +
  # scale_x_continuous(breaks = c(128, 256, 512, 1024, 2048, 4096)) + 
  scale_y_log10() +
  labs(y = 'community detection error rate', 
       colour = NULL) +
  geom_line(aes(x = n, y = med.err,
                colour = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q, ymax = third.q,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc,
                colour = 'SSC-ASE')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc, ymax = third.q.ssc,
                    colour = 'SSC-ASE'), width = .1) + 
  # geom_line(aes(x = n, y = med.err.ep,
  #               colour = 'MM-EP')) + 
  # geom_errorbar(aes(x = n, ymin = first.q.ep, ymax = third.q.ep,
  #                   colour = 'MM-EP'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc2,
                colour = 'SSC-A')) + 
  geom_errorbar(aes(x = n, ymin = first.q.ssc2, ymax = third.q.ssc2,
                    colour = 'SSC-A'), width = .1) + 
  # geom_line(aes(x = n, y = med.err.louvain,
  #               colour = 'MM-Louvain')) + 
  # geom_errorbar(aes(x = n, ymin = first.q.louvain, ymax = third.q.louvain,
  #                   colour = 'MM-Louvain'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```