### 2021-11-26 s675/rdpg.Rmd
notes on rdpg and grdpg

---
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: 
        - page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

set.seed(314159)
source('~/dev/pabm-grdpg/functions.R')
```

## {.plain}

\center

\LARGE

\textcolor{darkred}{Random Dot Product Graphs}

\normalsize

STAT-S 675

Fall 2021

# Recall From Last Time ...

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}
\newcommand{\argmin}{\mathrm{argmin}}

* Let $G = (V, E)$ be an undirected and hollow graph with $|V| = n$ 
and adjacency matrix $A$
  * $A \in \mathbb{R}^{n \times n}$ is symmetric with zero diagonals
* Suppose $G \sim F(\theta)$
  * What kind of $F(\theta)$ make sense here?
  * Given $F$ and observed $G$, how can we estimate $\theta$?

# Recall From Last Time ...

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

::: columns

:::: {.column width=55%}

$A_{ij} = \begin{cases} 
1 & \exists \text{ edge between } i \text{ and } j \\
0 & \text{else}
\end{cases}$

$A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.

\vspace*{1\baselineskip}

$A \sim \BG(P)$ iff:

1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.

For estimation, we need to impose some structure on $P$.

::::

:::: {.column width=45%}

**Example 1**: If every entry $P_{ij} = \theta \in (0, 1)$, 
then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
For this model, $A_{ij} \iid \Bernoulli(\theta)$.

```{r, fig.height = 2, fig.width = 2}
n <- 2 ** 5
p <- 1 / 5
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

::::

:::

# Random Dot Product Graph

Random Dot Product Graph $A \sim \RDPG(X)$  
(Young and Scheinerman, 2007)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
$x_i^\top x_j \in [0, 1]$
* $P = X X^\top \in \mathbb{R}^{n \times n}$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times d}$
* $A \sim \BG(P)$

* Estimation Objectives: 
  1. Estimate $X$ from $A$ (assume $d$ is fixed)
  2. If $x_1, ..., x_n \iid F(\theta)$, estimate $\theta$ from $A$
  (assuming certain $F$)

* Non-identifiability: For any $W \in \mathbb{O}(d)$, $X W$ is a latent configuration that produces the same $P = X X^\top = X W W^\top X^\top$

# Random Dot Product Graph

```{r}
n1 <- 50
n2 <- 50
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
u1 <- runif(n1)
u2 <- runif(n2)
x1 <- cos(pi / 3 * u1)
y1 <- sin(pi / 3 * u1)
x2 <- 1 + cos(pi / 3 * u2 + pi)
y2 <- 1 + sin(pi / 3 * u2 + pi)
data.matrix <- cbind(c(x1, x2), c(y1, y2))
```

::: columns

:::: {.column width=50%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent space')
```

::::

:::: {.column width=50%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), layout = 'spring', legend = FALSE)
title('RDPG')
```

::::

:::

# Maximum Likelihood Estimation

$$L(X | A) = \prod_{i < j} (x_i^\top x_j)^{A_{ij}} (1 - x_i^\top x_j)^{1 - A_{ij}}$$

* Intractable
* Not unique

# Adjacency Spectral Embedding

$$\hat{X} = \argmin_{X \in \mathbb{R}^{n \times d}} \|A - X X^\top \|_F^2$$

* Same solution as PCA
* $A = V \Lambda V^\top$
* $\hat{X} = V_d \Lambda_d^{1/2}$
  * $V_d \in \mathbb{R}^{n \times d}$
  * $\Lambda_d \in \mathbb{R}^{d \times d}$
* Not unique
  * If $\hat{X}$ is a solution, then so is $\hat{X} W$
  * Multiplying by $W$ preserves interpoint inner products and distances
* $\max_i \|\hat{x}_i - W x_i\| \stackrel{a.s.}{\to} 0$ (Athreya et al., 2018)

# Adjacency Spectral Embedding

::: columns

:::: {.column width=33%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
plot(data.matrix, col = z * 2, asp = 1, xlab = NA, ylab = NA,
     main = 'Latent space')
```

::::

:::: {.column width=33%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
P <- data.matrix %*% diag(c(1, 1)) %*% t(data.matrix)
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), layout = 'spring', legend = FALSE)
title('RDPG')
```

::::

:::: {.column width=33%}

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
par(mar = rep(1, 4))
A.eigen <- eigen(A)
Xhat <- A.eigen$vectors[, c(1, 2)] %*% diag(sqrt(abs(A.eigen$values[c(1, 2)])))
plot(Xhat, col = z * 2, asp = 1, main = 'ASE')
```

::::

:::

# Random Dot Product Graph

* For what types of graphs can we justify a latent structure?
* Any $A \sim \BG(P)$ is a RDPG if $P$ is positive semidefinite
* Whether we think of a graph as a RDPG depends on whether the latent structure is useful
* What about $P$ that are not positive semidefinite?

# Generalized Random Dot Product Graph

Generlized Random Dot Product Graph $A \sim \GRDPG(X)$  
(Rubin-Delanchy et al., 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p,q} x_j \in [0, 1]$
and $I_{p,q} = \blockdiag(I_p, -I_q)$
* $P = X I_{p,q} X^\top \in \mathbb{R}^{n \times n}$, 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top \in \mathbb{R}^{n \times (p+q)}$,  
$A \sim \BG(P)$

* Non-identifiability: For any $Q \in \mathbb{O}(p,q)$, $X Q$ is a latent configuration that produces the same 
$P = X I_{p,q} X^\top = X Q I_{p,q} Q^\top X^\top$
  * Multiplication by $Q$ **does not** preserve interpoint inner products or distances

* Any $A \sim \BG(P)$ is a GRDPG

# Adjacency Spectral Embedding for GRDPG

$$\hat{X} = \argmin_{X \in \mathbb{R}^{n \times (p+q)}} \|A - X I_{p,q} X^\top \|_F^2$$

* $A = V \Lambda V^\top$
* $\hat{X} = V_{p,q} |\Lambda_{p,q}|^{1/2}$
  * $V_{p,q} = \begin{bmatrix} V_{1:p} & | & V_{n-q+1:n} \end{bmatrix}$
  * $\Lambda_{p,q} = \diag(\lambda_1, ..., \lambda_p, \lambda_{n-q+1}, ..., \lambda_n)$
* Not unique
  * If $\hat{X}$ is a solution, then so is $\hat{X} Q$
* $\max_i \|\hat{x}_i - Q x_i \| \stackrel{a.s.}{\to} 0$ (Rubin-Delanchy et al., 2020)

# Next Time ...

* All $A \sim \BG(P)$ are GRDPGs
  * If $P$ is positive semidefinite, then it is a RDPG
* Includes SBM, DCBM, PABM
* What kind of latent structures correspond to block models?
* Can we use the ASE to estimate block model parameters?