### 2021-11-27 s675/blockmodels-and-rdpg.Rmd
notes for friday

---
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: 
        - page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

set.seed(314159)
source('~/dev/pabm-grdpg/functions.R')
```

## {.plain}

\center

\LARGE

\textcolor{darkred}{Block Models\\and\\(Generalized) Random Dot Product Graphs}

\normalsize

STAT-S 675

Fall 2021

# Recall From Last Time ...

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}
\newcommand{\argmin}{\mathrm{argmin}}

* Let $G = (V, E)$ be an undirected and hollow graph with $|V| = n$ 
and adjacency matrix $A$
  * $A \in \mathbb{R}^{n \times n}$ is symmetric with zero diagonals
* Suppose $G \sim F(\theta)$
  * What kind of $F(\theta)$ make sense here?
  * Given $F$ and observed $G$, how can we estimate $\theta$?

# Recall From Last Time ...

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

::: columns

:::: {.column width=55%}

$A_{ij} = \begin{cases} 
1 & \exists \text{ edge between } i \text{ and } j \\
0 & \text{else}
\end{cases}$

$A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.

\vspace*{1\baselineskip}

$A \sim \BG(P)$ iff:

1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.

For estimation, we need to impose some structure on $P$.

::::

:::: {.column width=45%}

**Example 1**: If every entry $P_{ij} = \theta \in (0, 1)$, 
then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
For this model, $A_{ij} \iid \Bernoulli(\theta)$.

```{r, fig.height = 2, fig.width = 2}
n <- 2 ** 5
p <- 1 / 5
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

::::

:::

# Recall From Last Time ...

Suppose each vertex $v_1, ..., v_n$ has hidden labels $z_1, ..., z_n \in [K]$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim \BG(P)$ is a *block model*.

**Example 2**: Stochastic Block Model with two communities

::: columns

:::: column

* $z_1, ..., z_n \in \{1, 2\}$
* $P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$

* To make this an assortative SBM, set $p q > r^2$.
* In this example, $p = 1/2$, $q = 1/4$, and $r = 1/8$.

::::

:::: column

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::

# Recall From Last Time ...

Erdos-Renyi Model (1959)

* $P_{ij} = \theta$ (not a block model)
* 1 parameter $\theta$

Stochastic Block Model (Lorrain and White, 1971)

* $P_{ij} = \theta_{z_i z_j}$
* $K (K + 1) / 2$ parameters $\theta_{kl}$

Degree Corrected Block Model (Karrer and Newman, 2011)

* $P_{ij} = \theta_{z_i z_j} \omega_i \omega_j$
* $K (K + 1) / 2 + n$ parameters $\theta_{kl}$, $\omega_i$

Popularity Adjusted Block Model (Sengupta and Chen, 2017)

* $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
* $K n$ parameters $\lambda_{ik}$

# Recall From Last Time ...

Random Dot Product Graph $A \sim \RDPG(X)$  
(Young and Scheinerman, 2007)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
$x_i^\top x_j \in [0, 1]$
* $A \sim \BG(X X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

Generalized Random Dot Product Graph $A \sim \GRDPG_{p, q}(X)$  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = \blockdiag(I_p, -I_q)$
* $A \sim \BG(X I_{p, q} X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

# Recall From Last Time ...

### Adjacency Spectral Embedding

To embed in $\mathbb{R}^d$,

1. Compute $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ 
where $\hat{\Lambda} \in \mathbb{R}^{d \times d}$ and 
$\hat{V} \in \mathbb{R}^{n \times d}$.  
For RDPG, use $d$ greatest eigenvalues; 
for GRDPG, use $p$ most positive and $q$ most negative eigenvalues.

2. For RDPG, let $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$; 
for GRDPG, let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.

\vspace*{.5\baselineskip}

RDPG: $\max\limits_i \|\hat{x}_i - W x_i \| \stackrel{a.s.}{\to} 0$
(Athreya et al., 2018)  
GRDPG: 
$\max\limits_i \|\hat{x}_i - Q x_i \| \stackrel{a.s.}{\to} 0$
(Rubin-Delanchy et al., 2020)

# Connecting Block Models to (G)RDPGs

All Bernoulli Graphs are RDPG (if $P$ is 
positive semidefinite)  
or GRDPG (in general).

**Example 2** (cont'd): Assortative SBM ($p q > r^2$) with $K = 2$

::: columns

:::: column

$$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$$

```{r fig.height = 3, fig.width = 6, out.width = '100%'}
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::: column

$$P = 
\begin{bmatrix} 
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} =
X X^\top$$

$$X = \begin{bmatrix} 
\sqrt{p} & 0 \\
\vdots & \vdots \\
\sqrt{p} & 0 \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
\vdots & \vdots \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
\end{bmatrix}$$

::::

:::

# Connecting the SBM to the RDPG

**Example 2** (cont'd): If we want to perform community detection,

1. Note that $A$ is a RDPG because $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
noting that as $n \to \infty$, the ASE approaches point masses.

```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

A.eigen <- eigen(A)
X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
plot(X.hat, asp = 1, col = z * 2, xlab = NA, ylab = NA, 
     main = 'ASE of the adjacency matrix drawn from SBM')
points(X, pch = 16, col = z * 2)
```

# Connecting the DCBM to the RDPG

**Example 3**: DCBM with two communities

::: columns

:::: column

* $z_1, ..., z_n \in \{1, 2\}$
* $P_{ij} = \begin{cases} 
p \omega_i \omega_j & z_i, z_j = 1 \\
q \omega_i \omega_j & z_i, z_j = 2 \\
r \omega_i \omega_j & z_i \neq z_j
\end{cases}$

* To make this an assortative DCBM, set $p q > r^2$.
* In this example, $p = q = 1/4$, $r = 1/2$, and $\omega_i \in (0, 1)$.

::::

:::: column

```{r, fig.height = 4, fig.width = 4, out.width = '100%'}
n1 <- 100
n2 <- 100
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/4
q <- 1/4
r <- 1/8
omega <- rbeta(n, 2, 1)
Omega <- tcrossprod(omega)
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
P <- P * Omega
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::

# Connecting the DCBM to the RDPG

**Example 3** (cont'd): Assortative DCBM ($p q > r^2$) with $K = 2$

::: columns

:::: column

$$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$$

```{r fig.height = 5, fig.width = 6, out.width = '100%'}
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::: column

$$P = 
\begin{bmatrix} 
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} =
X X^\top$$

$$X = \begin{bmatrix} 
\sqrt{p} \omega_1 & 0 \\
\vdots & \vdots \\
\sqrt{p} \omega_{n_1} & 0 \\
\sqrt{\frac{r^2}{p}} \omega_{n_1 + 1} & \sqrt{q - \frac{r^2}{p}} \omega_{n_1 + 1} \\
\vdots & \vdots \\
\sqrt{\frac{r^2}{p}} \omega_n & \sqrt{q - \frac{r^2}{p}} \omega_n
\end{bmatrix}$$

::::

:::

# Connecting the SBM to the RDPG

**Example 3** (cont'd): If we want to perform community detection,

1. Note that $A$ is a RDPG because $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply clustering algorithm on $\hat{X}$,  
noting that as $n \to \infty$, the ASE approaches line segments.

```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

A.eigen <- eigen(A)
X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
plot(X.hat, 
     # asp = 1, 
     col = z * 2, 
     xlab = NA, ylab = NA, 
     main = 'ASE of the adjacency matrix drawn from DCBM')
lines(rbind(c(0, 0), X[z == 1, ]), pch = 16, col = 2, lwd = 3)
lines(rbind(c(0, 0), X[z == 2, ]), pch = 16, col = 4, lwd = 3)
```

# Assortative SBM and DCBM are RDPGs

* $P$ is positive semidefinite $\implies$ RDPG
* Number of embedding dimensions $=$ number of communities
* Latent structure for the SBM are point masses,  
latent structure for the DCBM are line segments
* If the SBM or DCBM is not assortative, then $P$ is not positive semidefinite
  * GRDPG instead of RDPG
  * Number of embedding dimensions $=$ number of communities, 
  except split by positive and negative eigenvalues
  * ASE still converges to true latent positions up to multiplication by unidentifiable $Q \in \mathbb{O}(p, q)$
  * Multiplication by $Q$ doesn't change overall latent structure

# Demo

# Connecting the PABM to the GRDPG

**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.

**Lemma** (Noroozi, Rimal, and Pensky, 2020):

$A$ is sampled from a PABM if $P$ can be described as:

1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
between communities $k$ and $l$. 
2. Organize popularity parameters as vectors 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
of the $i$^th^ vertex of community $k$ towards community $l$. 
3. Each block can be decomposed as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.

**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.

# Connecting the PABM to the GRDPG

**Theorem**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim \GRDPG_{p, q}(X U)$ with

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$
* $U \in \mathbb{O}(K^2)$
* $X \in \mathbb{R}^{n \times K^2}$ is block diagonal and 
composed of $\{\lambda^{(kl)}\}_K$ 
with each block corresponding to a community.

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix} 
\in \mathbb{R}^{n \times K^2}$$

$$\Lambda^{(k)} = \begin{bmatrix} 
\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
\end{bmatrix} 
\in \mathbb{R}^{n_k \times K}$$

# Connecting the PABM to the GRDPG

::: columns

:::: column

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix}$$

::::

:::: column

\vspace*{2\baselineskip}

$$U \in \mathbb{O}(K^2)$$

::::

:::

$$A \sim \PABM(\{\lambda^{(kl)}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$

**Remark 1** (orthogonality of subspaces):
If $y_i^\top$ and $y_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $y_i^\top y_j = 0$.

**Remark 2** (non-uniqueness of the latent configuration):  
If $A \sim \GRDPG_{p, q}(Y)$, then $A \sim \GRDPG_{p, q}(Y Q)$ for any $Q$ in 
the indefinite orthogonal group with signature $p, q$.

**Remark 3**: Communities correspond to subspaces even with linear transformation $Q \in \mathbb{O}(p, q)$, but this may break the orthogonality property.

# Orthogonal Spectral Clustering

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$,  
then $B_{ij} = 0$ if $z_i \neq z_j$.

**Algorithm**: Orthogonal Spectral Clustering:

1. Let $V$ be the eigenvectors of $A$ corresponding to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs.

**Theorem** (KTT): 
Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\big((\log n)^{4c}\big)$,

$$
\max_{i, j} \hat{B}^{(ij)}_n = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
$$