### 2022-02-02 jsm/abstract/abstract.Rmd

---
title: Popularity Adjusted Block Models are Generalized Random Dot Product Graphs

# to produce blinded version set to 1
blinded: 0

authors: 
- name: John Koo
  affiliation: Department of Statistics, Indiana University

- name: Minh Tang
  affiliation: Department of Statistics, North Carolina State University

- name: Michael W. Trosset
  affiliation: Department of Statistics, Indiana University

keywords:
- network analysis, community detection, sparse subspace clustering, spectral clustering

abstract: |
  We connect two random graph models, the Popularity Adjusted Block Model 
  (PABM) and the Generalized Random Dot Product Graph (GRDPG), demonstrating
  that a PABM is a GRDPG in which communities correspond to certain mutually
  orthogonal subspaces of latent vectors. 
  This insight leads to the construction of new algorithms for community 
  detection and parameter estimation for the PABM, as well as improve an 
  existing algorithm that relies on Sparse Subspace Clustering. 
  Using established asymptotic properties of Adjacency Spectral Embedding for 
  the GRDPG, we derive asymptotic properties of these algorithms. 
  In particular, we demonstrate that the absolute number of community detection 
  errors tends to zero as the number of graph vertices tends to infinity. 
  Simulation experiments illustrate these properties.

bibliography: bibliography.bib
output: rticles::asa_article
---