### 2022-03-23 fls-lightning-talk/slides.Rmd
lightning talk slides

---
title: Popularity Adjusted Block Models are Generalized Random Dot Product Graphs
subtitle: Future Leaders Summit 2022 Lightning Presentation
author: John Koo, Indiana University
date: 'April 2022'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \usepackage{graphicx}
- \setbeamertemplate{itemize items}[circle]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

source('~/dev/pabm-grdpg/functions.R')
set.seed(314159)
```

## Contributors

::: columns

:::: {.column width='33%'}

```{r, out.width = '100px'}
knitr::include_graphics('john-koo.jpeg')
```

John Koo,  
PhD Student in Statistical Science,  
Indiana University

::::

:::: {.column width='33%'}

```{r, out.width = '100px'}
knitr::include_graphics('minh-tang.jpg')
```

Minh Tang,  
Assistant Professor of Statistics,  
NC State University

::::

:::: {.column width='33%'}

```{r, out.width = '100px'}
knitr::include_graphics('michael-trosset.jpg')
```

Michael Trosset,  
Professor of Statistics,  
Indiana University

::::

:::

## Community Detection for Networks

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}

```{r out.width = '50%', fig.height = 3, fig.width = 4}
Pz <- generate.P.beta(64, 4, 4, 1, 1, 4)
P <- Pz$P
z <- Pz$clustering
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
```

How can we cluster the nodes of a network?

Statistical inference (parametric approach):

1. Define a generative model for graph: 
$G \mid z_1, ..., z_n, \vec{\theta} \sim P(\vec{z}, \vec{\theta})$.
2. Develop a method for obtaining estimators: 
$f(G) = (\hat{\vec{z}}, \hat{\vec{\theta}})$.
3. Identify asymptotic properties of estimators: 
$(\hat{\vec{z}}, \hat{\vec{\theta}}) \to (\vec{z}, \vec{\theta})$.

## Overview

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

1. Block Models and the Popularity Adjusted Block Model

2. Generalized Random Dot Product Graphs

3. Connecting the PABM to the GRDPG

4. Community Detection for the PABM

# Block Models

## Bernoulli Graphs

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

::: columns

:::: {.column width=62.5%}

Let $G = (V, E)$ be an undirected and unweighted graph with $|V| = n$.

$G$ is described by adjacency matrix $A$ such that
$A_{ij} = \begin{cases} 
1 & \exists \text{ edge between } i \text{ and } j \\
0 & \text{else}
\end{cases}$

$A_{ji} = A_{ij}$ and $A_{ii} = 0$ $\forall i, j \in [n]$.

\vspace*{1\baselineskip}

$A \sim \BG(P)$ iff:

1. $P \in [0, 1]^{n \times n}$ describes edge probabilities between pairs of 
vertices.
2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.

::::

:::: {.column width=37.5%}

**Example 1**: If every entry $P_{ij} = \theta$, then $A \sim \BG(P)$ is an Erdos-Renyi graph.  
For this model, $A_{ij} \iid \Bernoulli(\theta)$.

```{r, fig.height = 2, fig.width = 2}
n <- 2 ** 5
p <- 1 / log(n)
P <- matrix(p, nrow = n, ncol = n)
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4)
```

::::

:::

## Block Models

Suppose each vertex $v_1, ..., v_n$ has labels $z_1, ..., z_n \in \{1, ..., K\}$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim \BG(P)$ is a *block model*.

**Example 2**: Stochastic Block Model with two communities

::: columns

:::: column

* $z_1, ..., z_n \in \{1, 2\}$
* $P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$

* To make this an assortative SBM, set $p q > r^2$.
* In this example, $p = 1/2$, $q = 1/4$, and $r = 1/8$.

::::

:::: column

```{r, fig.height = 3, fig.width = 4, out.width = '100%'}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::

## Block Models

Erdos-Renyi Model (1959)

* $P_{ij} = \theta$ (not a block model)
* 1 parameter $\theta$

Stochastic Block Model (Lorrain and White, 1971)

* $P_{ij} = \theta_{z_i z_j}$
* $K (K + 1) / 2$ parameters $\theta_{kl}$

Degree Corrected Block Model (Karrer and Newman, 2011)

* $P_{ij} = \theta_{z_i z_j} \omega_i \omega_j$
* $K (K + 1) / 2 + n$ parameters $\theta_{kl}$, $\omega_i$

Popularity Adjusted Block Model (Sengupta and Chen, 2017)

* $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
* $K n$ parameters $\lambda_{ik}$

## Popularity Adjusted Block Model

**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
Then $A \sim \PABM(P)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$,  
e.g., if $z_i = k$ and $z_j = l$, $P_{ij} = \lambda_{il} \lambda_{jk}$.

**Lemma** (Noroozi, Rimal, and Pensky, 2020):

$A$ is sampled from a PABM if $P$ can be described as:

1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
between communities $k$ and $l$. 
2. Organize popularity parameters as vectors 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
of the $i$^th^ vertex of community $k$ towards community $l$. 
3. Each block can be decomposed as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.

**Notation**: $A \sim \PABM(\{\lambda^{(kl)}\}_K)$.

# Generalized Random Dot Product Graphs

## (Generalized) Random Dot Product Graph Model

Random Dot Product Graph $A \sim \RDPG(X)$  
(Young and Scheinerman, 2007)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^d$ such that 
$x_i^\top x_j \in [0, 1]$
* $A \sim \BG(X X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

Generalized Random Dot Product Graph $A \sim \GRDPG_{p, q}(X)$  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

* Latent vectors $x_1, ..., x_n \in \mathbb{R}^{p+q}$ such that 
$x_i^\top I_{p, q} x_j \in [0, 1]$ and $I_{p, q} = \blockdiag(I_p, -I_q)$
* $A \sim \BG(X I_{p, q} X^\top)$ where 
$X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$

If latent vectors $X_1, ..., X_n \iid F$, then we write 
$(A, X) \sim \RDPG(F, n)$ or $(A, X) \sim \GRDPG_{p, q}(F, n)$.

## (Generalized) Random Dot Product Graph Model

### Recovery/Estimation

Want to estimate $X$ from $A$, or alternatively, 
interpoint distances, inner products, or angles.

### Adjacency Spectral Embedding

To embed in $\mathbb{R}^d$,

1. Compute $A \approx \hat{V} \hat{\Lambda} \hat{V}^\top$ 
where $\hat{\Lambda} \in \mathbb{R}^{d \times d}$ and 
$\hat{V} \in \mathbb{R}^{n \times d}$.  
For RDPG, use $d$ greatest eigenvalues; 
for GRDPG, use $p$ most positive and $q$ most negative eigenvalues.

2. For RDPG, let $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$; 
for GRDPG, let $\hat{X} = \hat{V} |\hat{\Lambda}|^{1/2}$.

\vspace*{.5\baselineskip}

RDPG: $\max\limits_i \|\hat{X}_i - W_n X_i \| \stackrel{a.s.}{\to} 0$
(Athreya et al., 2018)  
GRDPG: 
$\max\limits_i \|\hat{X}_i - Q_n X_i \| \stackrel{a.s.}{\to} 0$
(Rubin-Delanchy et al., 2020)

# Connecting the PABM to the GRDPG

## Connecting Block Models to the (G)RDPG Model

All Bernoulli Graphs are RDPG (if $P$ is 
positive semidefinite)  
or GRDPG (in general).

**Example 2** (cont'd): Assortative SBM ($p q > r^2$) with $K = 2$

::: columns

:::: column

$$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$$

```{r fig.height = 3, fig.width = 6, out.width = '100%'}
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::: column

$$P = 
\begin{bmatrix} 
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} =
X X^\top$$

$$X = \begin{bmatrix} 
\sqrt{p} & 0 \\
\vdots & \vdots \\
\sqrt{p} & 0 \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
\vdots & \vdots \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
\end{bmatrix}$$

::::

:::

## Connecting Block Models to the (G)RDPG Model

**Example 2** (cont'd): If we want to perform community detection,

1. Note that $A$ is a RDPG because $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
noting that as $n \to \infty$, the ASE approaches point masses.

```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

A.eigen <- eigen(A)
X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
plot(X.hat, asp = 1, col = z * 2, xlab = NA, ylab = NA, 
     main = 'ASE of the adjacency matrix drawn from SBM')
points(X, pch = 16, col = z * 2)
```

## Connecting Block Models to the (G)RDPG Model

::: columns

:::: column

```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
par(mar = rep(1.75, 4))
plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
```

\vspace*{.5\baselineskip}

```{r, out.width = '50%', fig.height = 2.5, fig.width = 2.5}
par(mar = rep(1.75, 4))
omega <- rbeta(n, 1, 1)
dc.matrix <- omega %*% t(omega)
P.dcbm <- P * dc.matrix
dcbm.eigen <- eigen(P.dcbm)
X.dcbm <- dcbm.eigen$vectors[, 1:2] %*% diag(dcbm.eigen$values[1:2] ** .5)
plot(X.dcbm, asp = 1, col = z, xlab = NA, ylab = NA, main = 'DCBM: Rays')

```

::::

:::: column

\vspace*{0\baselineskip}

```{r, out.width = '100%', fig.height = 4, fig.width = 4}
par(mar = rep(1.75, 4))
Pz <- generate.P.beta(n * 10)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P, 3, 1)
pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
```

::::

:::

## Connecting the PABM to the GRDPG

**Theorem** (KTT): $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim \GRDPG_{p, q}(X U)$ with

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$
* $U \in \mathbb{O}(K^2)$
* $X \in \mathbb{R}^{n \times K^2}$ is block diagonal and 
composed of $\{\lambda^{(kl)}\}_K$ 
with each block corresponding to a community.

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix} 
\in \mathbb{R}^{n \times K^2}$$

$$\Lambda^{(k)} = \begin{bmatrix} 
\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
\end{bmatrix} 
\in \mathbb{R}^{n_k \times K}$$

## Connecting the PABM to the GRDPG

::: columns

:::: column

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix}$$

::::

:::: column

\vspace*{2\baselineskip}

$$U \in \mathbb{O}(K^2)$$

::::

:::

$$A \sim \PABM(\{\lambda^{(kl)}\}_K) \iff A \sim \GRDPG_{p, q}(X U)$$

**Remark 1** (orthogonality of subspaces):
If $y_i^\top$ and $y_j^\top$ are two rows of $X U$ corresponding 
to different communities, then $y_i^\top y_j = 0$.

**Remark 2** (non-uniqueness of the latent configuration):  
If $A \sim \GRDPG_{p, q}(Y)$, then $A \sim \GRDPG_{p, q}(Y Q)$ for any $Q$ in 
the indefinite orthogonal group with signature $p, q$.

**Remark 3**: Communities correspond to subspaces even with linear transformation $Q \in \mathbb{O}(p, q)$, but this may break the orthogonality property.

# Community Detection for the PABM

## Orthogonal Spectral Clustering

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$,  
then $B_{ij} = 0$ if $z_i \neq z_j$.

**Algorithm**: Orthogonal Spectral Clustering:

1. Let $V$ be the eigenvectors of $A$ corresponding to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs.

**Theorem** (KTT): 
Let $\hat{B}_n$ with entries $\hat{B}_n^{(ij)}$ be the affinity matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\big((\log n)^{4c}\big)$,

$$
\max_{i, j} \hat{B}^{(ij)}_n = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)
$$

## Sparse Subspace Clustering

**Corollary**: The ASE of $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ lies near a collection of $K$-dimensional subspaces in $K^2$ dimensions.

**Algorithm**: Sparse Subspace Clustering (Elhamifar & Vidal, 2009):

1. Solve $n$ optimization problems $c_i = \arg\min_c \|c\|_1$ 
subject to $x_i = X^\top c$ and $c^{(i)} = 0$.  
This is typically performed via LASSO: 
$c_i = \arg\min \frac{1}{2} \|x_i - X_{-i}^\top c\|_2^2 + \lambda \|c\|_1$
2. Compile solutions $C = \begin{bmatrix} c_1 & \cdots & c_n \end{bmatrix}$.
3. Construct affinity matrix $B = |C| + |C^\top|$.

## Sparse Subspace Clustering

**Theorem** (KTT):

Let

* $P_n$ describe the edge probability matrix of the PABM with $n$ vertices, and 
$A_n \sim \BG(P_n)$;
* $\hat{V}_n$ be the matrix of eigenvectors of $A_n$ corresponding to the 
$K (K + 1) / 2$ most positive and $K (K - 1) / 2$ most negative eigenvalues.

Then

* For some $\lambda > 0$ and $N < \infty$, $\sqrt{n} \hat{V}_n$ obeys the Subspace Detection Property with probability 1 when $n > N$.

Remarks:

* For large $n$, we can identify $\lambda$ for SDP (Wang and Xu, 2016).
* SDP does not guarantee community detection.

## Simulation Results

```{r, fig.height = 3, fig.width = 8, out.width = '100%'}
clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-balanced.csv')

clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error.osc),
    first.q = quantile(error.osc, .25),
    third.q = quantile(error.osc, .75),
    med.err.ssc = median(error.ssc.ase),
    first.q.ssc = quantile(error.ssc.ase, .25),
    third.q.ssc = quantile(error.ssc.ase, .75),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75),
    med.err.ssc.A = median(error.ssc.A),
    first.q.ssc.A = quantile(error.ssc.A, .25),
    third.q.ssc.A = quantile(error.ssc.A, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  theme_bw() + 
  theme(legend.position = 'bottom') + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
                labels = c(expression(2^7), 
                           expression(2^8), 
                           expression(2^9), 
                           expression(2^10), 
                           expression(2^11), 
                           expression(2^12))) +
  scale_y_log10() + 
  labs(y = 'community detection error count', 
       colour = NULL, shape = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'OSC')) +
  geom_point(aes(x = n, y = med.err * n,
                 colour = 'OSC', shape = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC-ASE')) +
  geom_point(aes(x = n, y = med.err.ssc * n,
                 colour = 'SSC-ASE', shape = 'SSC-ASE')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC-ASE'), width = .1) +
  geom_line(aes(x = n, y = med.err.ssc.A * n,
                colour = 'SSC-A')) +
  geom_point(aes(x = n, y = med.err.ssc.A * n,
                 colour = 'SSC-A', shape = 'SSC-A')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
                    colour = 'SSC-A'), width = .1) +
  geom_line(aes(x = n, y = med.err.louvain * n,
                colour = 'MM-Louvain')) + 
  geom_point(aes(x = n, y = med.err.louvain * n,
                 colour = 'MM-Louvain', shape = 'MM-Louvain')) + 
  geom_errorbar(aes(x = n, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
                    colour = 'MM-Louvain'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```