### 2022-03-28 fls-lightning-talk/slides.Rmd
updated slides per discussion

---
title: Popularity Adjusted Block Models are Generalized Random Dot Product Graphs
subtitle: Future Leaders Summit 2022 Lightning Presentation
author: John Koo, Indiana University
date: 'April 2022'
output: 
  beamer_presentation:
    fig_crop: no
    theme: 'default'
    colortheme: 'beaver'
    includes:
      in_header: page_headers.tex
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{numbers,square,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \usepackage{graphicx}
- \setbeamertemplate{itemize items}[circle]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      comment = NA, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.align = 'center',
                      fig.lp = '')
options(xtable.comment = FALSE, 
        xtable.table.placement = 'H')
library(ggplot2)
import::from(magrittr, `%>%`)
theme_set(theme_bw())

source('~/dev/pabm-grdpg/functions.R')
set.seed(314159)
```

## Contributors

::: columns

:::: {.column width='33%'}

```{r, out.width = '100px'}
knitr::include_graphics('john-koo.jpeg')
```

John Koo,  
PhD Student in Statistical Science,  
Indiana University

::::

:::: {.column width='33%'}

```{r, out.width = '100px'}
knitr::include_graphics('minh-tang.jpg')
```

Minh Tang,  
Assistant Professor of Statistics,  
NC State University

::::

:::: {.column width='33%'}

```{r, out.width = '100px'}
knitr::include_graphics('michael-trosset.jpg')
```

Michael Trosset,  
Professor of Statistics,  
Indiana University

::::

:::

## Community Detection for Networks

\newcommand{\diag}{\text{diag}}
\newcommand{\tr}{\text{Tr}}
\newcommand{\blockdiag}{\text{blockdiag}}
\newcommand{\indep}{\stackrel{\text{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\text{iid}}{\sim}}
\newcommand{\Bernoulli}{\text{Bernoulli}}
\newcommand{\Betadist}{\text{Beta}}
\newcommand{\BG}{\text{BernoulliGraph}}
\newcommand{\Cat}{\text{Categorical}}
\newcommand{\Uniform}{\text{Uniform}}
\newcommand{\RDPG}{\text{RDPG}}
\newcommand{\GRDPG}{\text{GRDPG}}
\newcommand{\PABM}{\text{PABM}}

```{r out.width = '50%', fig.height = 3, fig.width = 4}
Pz <- generate.P.beta(64, 4, 4, 1, 1, 4)
P <- Pz$P
z <- Pz$clustering
A <- draw.graph(P)
qgraph::qgraph(A, groups = factor(z), legend = FALSE)
```

How might we cluster the nodes of a network?

1. Define a probability model with communities that might generate the graph (e.g., popularity adjusted block model).
2. Develop estimators for the parameters of the probability model, including the community labels.
3. Describe the properties of the estimators (e.g., consistency).

## Bernoulli Graphs

<style type="text/css">
.caption {
    font-size: x-small;
}
</style>

Let $G$ be an undirected and unweighted graph with $n$ vertices.

$G$ is described by adjacency matrix $A$ such that
$A_{ij} = \begin{cases} 
1 & \text{an edge connects vertices } i \text{ and } j \\
0 & \text{otherwise}
\end{cases}$

$A_{ji} = A_{ij}$ and $A_{ii} = 0$.

\vspace*{1\baselineskip}

$A \sim \BG(P)$ iff:

1. $P$ is an $n \times n$ matrix of edge probabilities between pairs of vertices.
2. $A_{ij} \indep \Bernoulli(P_{ij})$ for each $i < j$.

## Block Models

Suppose each vertex $v_1, ..., v_n$ has labels $z_1, ..., z_n \in \{1, ..., K\}$,  
and each $P_{ij}$ depends on labels $z_i$ and $z_j$.  
Then $A \sim \BG(P)$ is a *block model*.

**Example 1**: Stochastic Block Model with $K=2$ communities.

::: columns

:::: column

$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$

::::

:::: column

```{r, fig.height = 3, fig.width = 4, out.width = '100%', fig.cap = 'Two community SBM with $p=1/2$, $q=1/4$, $r=1/8$'}
n1 <- 2 ** 5
n2 <- 2 ** 5
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/4
r <- 1/8
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::

## Popularity Adjusted Block Model

**Def** Popularity Adjusted Block Model (Sengupta and Chen, 2017):

Let each vertex $i \in [n]$ have $K$ popularity parameters
$\lambda_{i1}, ..., \lambda_{iK} \in [0, 1]$. 
Then $A \sim \PABM(\{\lambda_{ik}\}_K)$ if each $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$.

**Def** (Noroozi, Rimal, and Pensky, 2020):

$A$ is sampled from a PABM if $P$ can be described as:

1. Let each $P^{(kl)}$ denote the $n_k \times n_l$ matrix of edge probabilities 
between communities $k$ and $l$. 
2. Organize popularity parameters as vectors 
$\lambda^{(kl)} \in \mathbb{R}^{n_k}$ 
such that $\lambda^{(kl)}_i = \lambda_{k_i l}$ is the popularity parameter 
of the $i$^th^ vertex of community $k$ towards community $l$. 
3. Each block can be decomposed as 
$P^{(kl)} = \lambda^{(kl)} (\lambda^{(lk)})^\top$.

## Generalized Random Dot Product Graph

**Def** Generalized Random Dot Product Graph  
(Rubin-Delanchy, Cape, Tang, Priebe, 2020)

Let $I_{p,q} = \blockdiag(I_p, -I_q)$ and suppose that $x_1, \ldots, x_n \in \mathbb{R}^{p+q}$ are such that $x_i^\top I_{p,q} x_j \in [0,1]$.  
Then $A \sim \GRDPG_{p, q}(X)$ iff $A \sim \BG(X I_{p,q} X^\top)$, where $X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top$.

Adjacency Spectral Embedding (Sussman et al., 2012) estimates $x_1, ..., x_n \in \mathbb{R}^{p+q}$ from $A$:

1. Let $\hat{\Lambda}$ be the diagonal matrix that contains the absolute values of the $p$ most positive and the $q$ most negative eigenvalues.
2. Let $\hat{V}$ be the matrix whose columns are the corresponding eigenvectors.
3. Compute $\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.

**Theorem**: $\max\limits_i \|\hat{X}_i - Q_n X_i \| = O_P \Big( \frac{(\log n)^c}{n^{1/2}} \Big)$ 
as $n \to \infty$

## Connecting Block Models to the GRDPG Model

All Bernoulli Graphs are GRDPGs.

**Example 1** (cont'd): SBM with $K = 2$.

::: columns

:::: column

$$P_{ij} = \begin{cases} 
p & z_i = z_j = 1 \\
q & z_i = z_j = 2 \\
r & z_i \neq z_j
\end{cases}$$

```{r fig.height = 3, fig.width = 6, out.width = '100%'}
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

::::

:::: column

$$P = 
\begin{bmatrix} 
P^{(11)} & P^{(12)} \\
P^{(21)} & P^{(22)}
\end{bmatrix} =
X I_{2,0} X^\top$$

$$X = \begin{bmatrix} 
\sqrt{p} & 0 \\
\vdots & \vdots \\
\sqrt{p} & 0 \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p} \\
\vdots & \vdots \\
\sqrt{r^2 / p} & \sqrt{q - r^2 / p}
\end{bmatrix}$$

::::

:::

## Connecting Block Models to the GRDPG Model

**Example 1** (cont'd): To perform community detection,

1. Note that $A$ is a RDPG because $P = X X^\top$.
2. Compute the ASE $A \approx \hat{X} \hat{X}^\top$ with 
$\hat{X} = \hat{V} \hat{\Lambda}^{1/2}$.
3. Apply a clustering algorithm (e.g., $K$-means) to $\hat{X}$,  
noting that $\hat{X}$ approaches point masses as $n \to \infty$.

```{r, fig.height = 5, fig.width = 5, out.width = '50%'}
P.eigen = eigen(P)
X <- P.eigen$vectors[, 1:2] %*% diag(P.eigen$values[1:2] ** .5)

A.eigen <- eigen(A)
X.hat <- A.eigen$vectors[, 1:2] %*% diag(A.eigen$values[1:2] ** .5)
plot(X.hat, asp = 1, col = z * 2, xlab = NA, ylab = NA, 
     main = 'ASE of the adjacency matrix drawn from SBM')
points(X, pch = 16, col = z * 2)
```

## Connecting Block Models to the GRDPG Model

::: columns

:::: column

```{r, out.width = '100%', fig.height = 4, fig.width = 4}
# par(mar = rep(1.75, 4))
plot(X, asp = 1, col = z, xlab = NA, ylab = NA, main = 'SBM: Point masses')
```

::::

:::: column

\vspace*{0\baselineskip}

```{r, out.width = '100%', fig.height = 4, fig.width = 4}
par(mar = rep(1.75, 4))
Pz <- generate.P.beta(n * 10)
P <- Pz$P
z <- Pz$clustering
X <- embedding(P, 3, 1)
pairs(X, col = z, asp = 1, pch = '.', main = 'PABM: Orthogonal subspaces')
```

::::

:::

## Connecting the PABM to the GRDPG

**Theorem** (KTT): $A \sim \PABM(\{\lambda^{(kl)}\}_K)$ is equivalent to 
$A \sim \GRDPG_{p, q}(X U)$ with

* $p = K (K + 1) / 2$, $q = K (K - 1) / 2$;
* $U$ is an orthogonal matrix;
* $X \in \mathbb{R}^{n \times K^2}$ is a block diagonal matrix 
composed of popularity vectors with each block corresponding to a community.

$$X = \begin{bmatrix}
\Lambda^{(1)} & \cdots & 0 \\
0 & \ddots & 0 \\
0 & \cdots & \Lambda^{(K)}
\end{bmatrix} 
\in \mathbb{R}^{n \times K^2}$$

$$\Lambda^{(k)} = \begin{bmatrix} 
\lambda^{(k1)} & \cdots & \lambda^{(kK)} 
\end{bmatrix} 
\in \mathbb{R}^{n_k \times K}$$

$$A \sim \PABM(\{\lambda_{ik}\}_K) \text{ iff } A \sim \GRDPG_{p, q}(X U)$$

## Orthogonal Spectral Clustering

**Theorem** (KTT): 
If $P = V \Lambda V^\top$ and $B = n V V^\top$,  
then $B_{ij} = 0$ if $z_i \neq z_j$.

**Algorithm**: Orthogonal Spectral Clustering:

1. Let $V$ be the eigenvectors of $A$ corresponding to the $K (K+1)/2$ most 
positive and $K (K-1) / 2$ most negative eigenvalues.
2. Compute $B = |n V V^\top|$ applying $|\cdot|$ entry-wise.
3. Construct graph $G$ using $B$ as its similarity matrix.
4. Partition $G$ into $K$ disconnected subgraphs.

**Theorem** (KTT): 
Let $\hat{B}$ with entries $\hat{B}_{ij}$ be the affinity matrix from OSC. 
Then $\forall$ pairs $(i, j)$ belonging to different communities 
and sparsity factor satisfying $n \rho_n = \omega\big((\log n)^{4c}\big)$,

$\max_{i, j} \hat{B}_{ij} = 
O_P \Big( \frac{(\log n)^c}{\sqrt{n \rho_n}} \Big)$ 
as $n \to \infty$.

## Simulation Results

We compared a variety of algorithms for community detection on randomly generated PABMs:

* Modularity Maximization (Sengupta and Chen) using the Louvain algorithm;
* Orthogonal Spectral Clustering (KTT);
* Sparse Subspace Clustering on the ASE (KTT);
* Sparse Subspace Clustering on the columns of $A$ (Noorozi, Rimal, Pensky).

```{r, fig.height = 3, fig.width = 8, out.width = '100%'}
clustering.df <- readr::read_csv('~/dev/pabm-grdpg/simulation-results/cluster-sim-balanced.csv')

clustering.df %>%
  dplyr::group_by(n, K) %>%
  dplyr::summarise(
    med.err = median(error.osc),
    first.q = quantile(error.osc, .25),
    third.q = quantile(error.osc, .75),
    med.err.ssc = median(error.ssc.ase),
    first.q.ssc = quantile(error.ssc.ase, .25),
    third.q.ssc = quantile(error.ssc.ase, .75),
    med.err.louvain = median(error.louvain),
    first.q.louvain = quantile(error.louvain, .25),
    third.q.louvain = quantile(error.louvain, .75),
    med.err.ssc.A = median(error.ssc.A),
    first.q.ssc.A = quantile(error.ssc.A, .25),
    third.q.ssc.A = quantile(error.ssc.A, .75)
  ) %>% 
  dplyr::ungroup() %>% 
  ggplot() +
  theme_bw() + 
  theme(legend.position = 'bottom') + 
  scale_x_log10(breaks = c(128, 256, 512, 1024, 2048, 4096),
                labels = c(expression(2^7), 
                           expression(2^8), 
                           expression(2^9), 
                           expression(2^10), 
                           expression(2^11), 
                           expression(2^12))) +
  scale_y_log10() + 
  labs(y = 'community detection error count', 
       colour = NULL, shape = NULL) +
  geom_line(aes(x = n, y = med.err * n,
                colour = 'OSC')) +
  geom_point(aes(x = n, y = med.err * n,
                 colour = 'OSC', shape = 'OSC')) +
  geom_errorbar(aes(x = n, ymin = first.q * n, ymax = third.q * n,
                    colour = 'OSC'), width = .1) + 
  geom_line(aes(x = n, y = med.err.ssc * n,
                colour = 'SSC-ASE')) +
  geom_point(aes(x = n, y = med.err.ssc * n,
                 colour = 'SSC-ASE', shape = 'SSC-ASE')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc * n, ymax = third.q.ssc * n,
                    colour = 'SSC-ASE'), width = .1) +
  geom_line(aes(x = n, y = med.err.ssc.A * n,
                colour = 'SSC-A')) +
  geom_point(aes(x = n, y = med.err.ssc.A * n,
                 colour = 'SSC-A', shape = 'SSC-A')) +
  geom_errorbar(aes(x = n, ymin = first.q.ssc.A * n, ymax = third.q.ssc.A * n,
                    colour = 'SSC-A'), width = .1) +
  geom_line(aes(x = n, y = med.err.louvain * n,
                colour = 'MM-Louvain')) + 
  geom_point(aes(x = n, y = med.err.louvain * n,
                 colour = 'MM-Louvain', shape = 'MM-Louvain')) + 
  geom_errorbar(aes(x = n, ymin = first.q.louvain * n, ymax = third.q.louvain * n,
                    colour = 'MM-Louvain'), width = .1) + 
  scale_colour_brewer(palette = 'Set1') + 
  facet_wrap(~ K, labeller = 'label_both')
```

## Simulation Setup

1. $z_1, ..., z_n \iid \Cat(1/K, ..., 1/K)$
2. $\lambda_{ik} \iid \Betadist(a_{ik}, b_{ik})$  
    * $a_{ik} = \begin{cases} 2 & z_i = k \\ 1 & z_i \neq k \end{cases}$   
    * $b_{ik} = \begin{cases} 1 & z_i = k \\ 2 & z_i \neq k \end{cases}$
3. $P_{ij} = \lambda_{i z_j} \lambda_{j z_i}$
4. $A \sim \BG(P)$